---
title: "OATLD Data Analysis"
output: html_notebook
---

The purpose of this project is to look at ASVs that are elevated in OATLD-afflicted corals.

```{r}
#Load the libraries - NEEDS TO BE CLEANED UP
library(phyloseq)
library(metagMisc)
library(nationalparkcolors)
library(phyloseq)
library(microbiome)
library(CoDaSeq)
library(zCompositions)
library(cowplot)
library(googleway)
library(ggplot2)
library(ggrepel)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(vegan)
library(tidyverse)
library(reshape2)
library(stringr)
library(RColorBrewer)
library(corncob)
library(DESeq2)
library(Biostrings)
library(seqateurs)
library(DECIPHER)
library(phangorn)
library(seqinr)
library(ape)
library(ggtree)
library(purrr)
library(digest)
library(decontam)
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```

```{r}
#Create the phyloseq object
otu <- read.table("silva_nochloronomito_otu_table.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)

OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps
summarize_phyloseq(ps)

read_counts <- rowSums(otu_table(ps))
View(read_counts)
read_counts<- as.data.frame(read_counts)
```

```{r}
#Get relative abundance
ps_ra<- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
```

Filter the dataset

```{r}
#filter based on prevalence and max relative abundance
#Prevalence of a given ASV
prevdf = apply(X = otu_table(ps),
               MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
#maximum relative abundance of each ASV
relabun = apply(otu_table(ps_ra), 
                MARGIN = ifelse(taxa_are_rows(ps_ra), yes = 1, no = 2),
                max)

#create prev and relabun data table w/ taxonomy
prevdf = data.frame(Prevalence = prevdf,
                    MaxRelAbundance = relabun,
                    tax_table(ps))


# Execute rel abun and prev filter, using `prune_taxa()` function
# Define abundance and prevelence threshold for taxa 
# must be >= 0.1% in at least 1 sample, and be in 3 or more samples
relAbundanceThreshold <- 0.001
prevalenceThreshold <- 3

keepTaxa <- rownames(prevdf)[(prevdf$MaxRelAbundance >= relAbundanceThreshold &
                                prevdf$Prevalence >= prevalenceThreshold)]

ps_f1 <- prune_taxa(keepTaxa, ps) #ps filt obj

#Write out filtered Dataset Table
otu_ps_f1 = as(otu_table(ps_f1), "matrix")
taxon_ps_f1 = as(tax_table(ps_f1), "matrix")
write.table(otu_ps_f1,"silva_nochloronomito_otu_table_ps_f1.txt",sep="\t",col.names=NA)
write.table(taxon_ps_f1,"silva_nochloronomito_taxa_table_ps_f1.txt",sep="\t",col.names=NA)


```

```{r}
#Beta Diversity Analyses
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu_ps_f1), method="CZM", label=0, z.warning = 0.999)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)

```

```{r}
#Set colors for palette
palette1<-c("blank"= "#26547c", "DD"= "#ef476f", "HD" ="#ffd166", "HH" = "#06d6a0")

#Plot Beta diversity PCA
pdf("PCA_Condition.pdf",bg = "white",width=8.5)
p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=samples$Condition, shape=samples$Site))
p<-p+geom_point( size=3.5)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(legend.position = "right")+
  theme(legend.title = element_blank()) +
  theme(legend.key.width = unit(1.0,  unit = "cm")) +
  guides(fill = guide_legend(byrow = TRUE)) 
p<- p+ labs(x=xlab, y=ylab, fill="Condition") + coord_fixed()+
  labs(fill = "Condition", shape="Site")+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_fill_manual(values = palette1) +
  theme_minimal()
p
dev.off()
```

First, remove sequences from the blanks using the decontamination package.

```{r}
#What Sequences/ASVs Are Actually In the Blanks?
ps_blanks<- subset_samples(ps, Condition =="blank")
ps_blanks_pruned <- prune_taxa(taxa_sums(ps_blanks) > 0, ps_blanks)

ps_blanks_pruned
View(ps_blanks_pruned@tax_table)
write.csv(ps_blanks_pruned@tax_table, "ASVs_in_blanks.csv")

#Visualize library size
df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Condition)) + geom_point()

sample_data(ps)$is.neg <- sample_data(ps)$Condition == "blank"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg")
table(contamdf.prev$contaminant)
head(which(contamdf.prev$contaminant))

contamdf.prev05 <- isContaminant(ps, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant)

contam_list<- contamdf.prev05[contamdf.prev05$contaminant=="TRUE",]
contam_list_w_tax<- merge(contam_list, taxon, by = "row.names")

write.csv(contam_list_w_tax, "true_contaminants.csv")

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Condition == "blank", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Condition == "DD" | sample_data(ps.pa)$Condition == "HD" | sample_data(ps.pa)$Condition == "HH", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")

#Filter out Contaminants
ps_no_contam <- prune_taxa(!contamdf.prev$contaminant, ps)
ps_no_contam <- subset_samples(ps_no_contam, Condition != "blank")

ps_no_contam_ra<- transform_sample_counts(ps_no_contam, function(OTU) OTU/sum(OTU))

summarize_phyloseq(ps_no_contam)
ps_no_contam

```

Try beta diversity metrics again now that contaminating sequences and blanks are removed

```{r}
#filter based on prevalence and max relative abundance
#Prevalence of a given ASV
prevdf = apply(X = otu_table(ps_no_contam),
               MARGIN = ifelse(taxa_are_rows(ps_no_contam), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
#maximum relative abundance of each ASV
relabun = apply(otu_table(ps_no_contam_ra), 
                MARGIN = ifelse(taxa_are_rows(ps_no_contam_ra), yes = 1, no = 2),
                max)

#create prev and relabun data table w/ taxonomy
prevdf = data.frame(Prevalence = prevdf,
                    MaxRelAbundance = relabun,
                    tax_table(ps_no_contam))


# Execute rel abun and prev filter, using `prune_taxa()` function
# Define abundance and prevelence threshold for taxa 
# must be >= 0.1% in at least 1 sample, and be in 2 or more samples
relAbundanceThreshold <- 0.001
prevalenceThreshold <- 2

keepTaxa <- rownames(prevdf)[(prevdf$MaxRelAbundance >= relAbundanceThreshold &
                                prevdf$Prevalence >= prevalenceThreshold)]

ps_f2 <- prune_taxa(keepTaxa, ps_no_contam) #ps filt obj

#Write out filtered Dataset Table
otu_ps_f2 = as(otu_table(ps_f2), "matrix")
taxon_ps_f2 = as(tax_table(ps_f2), "matrix")
samples_ps_f2<- as(sample_data(ps_f2), "matrix")
write.table(otu_ps_f2,"silva_nochloronomito_otu_table_ps_f2.txt",sep="\t",col.names=NA)
write.table(taxon_ps_f2,"silva_nochloronomito_taxa_table_ps_f2.txt",sep="\t",col.names=NA)

ps_f2

ps_f2_ra<- transform_sample_counts(ps_f2, function(OTU) OTU/sum(OTU))
```

```{r}
#Beta Diversity Analyses
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu_ps_f2), method="CZM", label=0, z.warning = 0.999)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)
```

```{r}
palette2<- c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")

metadata_ps_f2<- as.data.frame(samples_ps_f2)

p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=metadata_ps_f2$Condition, shape=metadata_ps_f2$Site))
p<-p+geom_point( size=3.5)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(legend.position = "right")+
  theme(legend.title = element_blank()) +
  theme(legend.key.width = unit(1.0,  unit = "cm")) +
  guides(fill = guide_legend(byrow = TRUE)) 
p<- p+ labs(x=xlab, y=ylab, fill="Condition") + coord_fixed()+
  labs(fill = "Condition", shape="Site")+
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_fill_manual(values = palette2, breaks = c("HH", "HD", "DD"), labels = c("H", "U", "D")) +
  theme_minimal() +
   labs(x = "Principal Component 1: 30%", y = "Principal Component 2: 5%") +
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black")) +
  guides(fill = guide_legend(override.aes = list(shape = 21)))
p

pdf("beta_diversity_PCA_health_site.pdf", width = 8, height=6)
p
dev.off()
```

```{r}
#Run the PERMANOVA w Site and Condition
###################PERMANOVA#################
dist.clr <- dist(E.clr)
perm<-adonis2(dist.clr~Condition,as(sample_data(ps_f2),"data.frame"))
print(perm)

write.table(perm, "permanova_results_Condition.txt")

perm<-adonis2(dist.clr~Site,as(sample_data(ps_f2),"data.frame"))
print(perm)

write.table(perm, "permanova_results_Site.txt")
```

Because health condition is a significant factor, we need to run the pairwise PERMANOVA to determine which conditions are different from each other. First, we must read in the pairwise.adonis() function. Then we will run it.

```{r}
#Here we are going to try to subset the different locations and re-run the PERMANOVA
pairwise.adonis <- function(x,factors, sim.function = 'vegdist', sim.method = 'bray', p.adjust.m ='bonferroni',reduce=NULL,perm=999)
{

  co <- combn(unique(as.character(factors)),2)
  pairs <- c()
  Df <- c()
  SumsOfSqs <- c()
  F.Model <- c()
  R2 <- c()
  p.value <- c()


  for(elem in 1:ncol(co)){
    if(inherits(x, 'dist')){
      x1=as.matrix(x)[factors %in% c(as.character(co[1,elem]),as.character(co[2,elem])),
                      factors %in% c(as.character(co[1,elem]),as.character(co[2,elem]))]
      }

    else  (
      if (sim.function == 'daisy'){
            x1 = daisy(x[factors %in% c(co[1,elem],co[2,elem]),],metric=sim.method)
        }
      else{x1 = vegdist(x[factors %in% c(co[1,elem],co[2,elem]),],method=sim.method)}
    )

    x2 = data.frame(Fac = factors[factors %in% c(co[1,elem],co[2,elem])])

    ad <- adonis2(x1 ~ Fac, data = x2,
                 permutations = perm);
    pairs <- c(pairs,paste(co[1,elem],'vs',co[2,elem]));
    Df <- c(Df,ad$Df[1])
    SumsOfSqs <- c(SumsOfSqs,ad$SumOfSqs[1])
    F.Model <- c(F.Model,ad$F[1]);
    R2 <- c(R2,ad$R2[1]);
    p.value <- c(p.value,ad$`Pr(>F)`[1])
  }
  p.adjusted <- p.adjust(p.value,method=p.adjust.m)

  sig = c(rep('',length(p.adjusted)))
  sig[p.adjusted <= 0.05] <-'.'
  sig[p.adjusted <= 0.01] <-'*'
  sig[p.adjusted <= 0.001] <-'**'
  sig[p.adjusted <= 0.0001] <-'***'
  pairw.res <- data.frame(pairs,Df,SumsOfSqs,F.Model,R2,p.value,p.adjusted,sig)

  if(!is.null(reduce)){
    pairw.res <- subset (pairw.res, grepl(reduce,pairs))
    pairw.res$p.adjusted <- p.adjust(pairw.res$p.value,method=p.adjust.m)

    sig = c(rep('',length(pairw.res$p.adjusted)))
    sig[pairw.res$p.adjusted <= 0.1] <-'.'
    sig[pairw.res$p.adjusted <= 0.05] <-'*'
    sig[pairw.res$p.adjusted <= 0.01] <-'**'
    sig[pairw.res$p.adjusted <= 0.001] <-'***'
    pairw.res <- data.frame(pairw.res[,1:7],sig)
  }
  class(pairw.res) <- c("pwadonis", "data.frame")
  return(pairw.res)
}

### Method summary
summary.pwadonis = function(object, ...) {
  cat("Result of pairwise.adonis:\n")
  cat("\n")
  print(object, ...)
  cat("\n")
  cat("Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1\n")
}

set.seed(123)
## Run pairwise permanova
groups <- as.character(sample_data(ps_f2)$Condition) # Apparently, my function doesn't like factors
pairwise.adonis(x=dist.clr, factors= groups)
```

```{r}
## Run pairwise permanova
groups <- as.character(sample_data(ps_f2)$Site) # Apparently, my function doesn't like factors
pairwise.adonis(x=dist.clr, factors= groups)
```

```{r}
### Beta Diversity Dispersions - Condition
#calculate multivariate dispersions based on Condition
mod <-betadisper(dist.clr, metadata_ps_f2$Condition)
#one way anova
anova(mod)
#boxplots
pdf("beta_dispersion_condition.pdf",width=8.5)
plot(mod)
dev.off()
```

```{r}
## Compute mean distance to centroid per group
#this just prints values on the console
tapply(mod$distances, metadata_ps_f2$Condition, mean)
## Same, but variance instead
tapply(mod$distances, metadata_ps_f2$Condition, var)

#Get the distances to centroid from the model
mod$distances
dis <- mod$distances
#melt
dis.melt <- melt(dis)
#move rownames to columns so we can merge the dispersion values and metadata
dis.melt$Sample <- rownames(dis.melt)
samples$Sample <- rownames(samples)
#merge metadata and dispersion 
dis.condition <- merge(samples, dis.melt)
#rename column
colnames(dis.condition)[6] <- "distance"

shapiro.test(dis.condition$distance) #Not normally distributed, use kruskal test
distance.test<- kruskal.test(distance ~ Condition, data=dis.condition)
print(distance.test)

capture.output(distance.test,file="kruskal_distance_to_centroid.txt")

#Post Hoc Test
library(FSA)
dunnTest(distance ~ Condition,
  data = dis.condition,
  method = "bonferroni"
)
```

```{r}
library(ggpubr)
p2<-ggplot(dis.condition, aes(x=Condition, y=distance)) +
  geom_violin(mapping =aes(x=Condition,y=distance, fill=Condition, color="black"), alpha = 0.6) +
  geom_point(position=position_jitter(width=.1, height=0),aes(colour="black",fill=Condition),size=4, pch=21, colour = "black") +
  scale_fill_manual(values = palette2) +
  scale_color_manual(values = c( "black")) +
  ylab("Distance to Centroid")+
  xlab("Health Condition")+
  guides(color="none") +
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(colour="black"))+
  theme(axis.text.y = element_text(colour="black"))+
  theme(axis.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))+
  theme(axis.title.x = element_text(size = 16))+
  theme(axis.title.y = element_text(size = 16))+
  theme(strip.text.x=element_text(size=16, face="bold"))+
  theme(legend.position = "none") +
  stat_compare_means(
    comparisons = list(c("DD", "HH")),
    method = "t.test",
    label = "p.signif")
p2
```

```{r}
#Make it a boxplot
library(ggpubr)
p2<-ggplot(dis.condition, aes(x=factor(Condition, levels=c("HH", "HD", "DD")), y=distance)) +
  geom_boxplot(mapping =aes(x=Condition,y=distance, fill=Condition, color="black"), alpha = 0.6) +
  geom_point(position=position_jitter(width=.1, height=0),aes(colour="black",fill=Condition),size=4, pch=21, colour = "black") +
  scale_fill_manual(values = palette2) +
  scale_color_manual(values = c( "black")) +
  ylab("Distance to Centroid")+
  xlab("Health Condition")+
  guides(color="none") +
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(colour="black"))+
  theme(axis.text.y = element_text(colour="black"))+
  theme(axis.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))+
  theme(axis.title.x = element_text(size = 16))+
  theme(axis.title.y = element_text(size = 16))+
  theme(strip.text.x=element_text(size=16, face="bold"))+
  theme(legend.position = "none") +
  stat_compare_means(
    comparisons = list(c("DD", "HH")),
    method = "t.test",
    label = "p.signif") +
  scale_x_discrete(labels = c("DD" = "HH", 
                              "HD" = "HD",
                              "HH" = "DD"))
p2
```

```{r}
#What Microbes are present?
library(metacoder)

set.seed(123)

obj <- parse_phyloseq(ps_f2)

obj %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Class", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_plot_class.pdf")

obj %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_plot_order.pdf")
```

```{r}
ps_f2_HH<- subset_samples(ps_f2, Condition =="HH")
ps_f2_HH_pruned <- prune_taxa(taxa_sums(ps_f2_HH) > 0, ps_f2_HH)


obj_HH <- parse_phyloseq(ps_f2_HH_pruned)

obj_HH %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Class", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_HH_plot_class.pdf")

obj_HH %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_HH_plot_order.pdf")

```

```{r}
ps_f2_HD<- subset_samples(ps_f2, Condition =="HD")
ps_f2_HD_pruned <- prune_taxa(taxa_sums(ps_f2_HD) > 0, ps_f2_HD)

obj_HD <- parse_phyloseq(ps_f2_HD_pruned)

obj_HD %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Class", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_HD_plot_class.pdf")

obj_HD %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_HD_plot_order.pdf")

```

```{r}
ps_f2_DD<- subset_samples(ps_f2, Condition =="DD")
ps_f2_DD_pruned <- prune_taxa(taxa_sums(ps_f2_DD) > 0, ps_f2_DD)

ps_f2_DD_pruned

obj_DD <- parse_phyloseq(ps_f2_DD_pruned)

obj_DD %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Class", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_DD_plot_class.pdf")

obj_DD %>%
  filter_taxa(grepl(pattern = "^[a-zA-Z]+$", taxon_names)) %>% # remove "odd" taxa
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = n_obs,
            node_color = n_obs,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_DD_plot_order.pdf")

```

```{r}
#Try it again with relative abundance
obj_DD = parse_phyloseq(ps_f2_DD_pruned)

print(obj_DD$data$sample_data$sample_id)
print(obj_DD$data$tax_data$otu_id)
print(colnames(obj_DD$data$otu_table))

obj_DD$data$otu_props <- calc_obs_props(obj_DD, "otu_table", cols = obj_DD$data$sample_data$sample_id)

obj_DD$data$tax_abund <- calc_taxon_abund(obj_DD, "otu_props", cols = obj_DD$data$sample_data$sample_id)

obj_DD$data$tax_mean_prop <- calc_group_mean(obj_DD, data = "tax_abund",
                                          cols = obj_DD$data$sample_data$sample_id,
                                          groups = "mean_prop")

obj_DD %>% 
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = mean_prop,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_DD_plot_order_RA.pdf")

```

```{r}
obj_HD = parse_phyloseq(ps_f2_HD_pruned)

print(obj_HD$data$sample_data$sample_id)
print(obj_HD$data$tax_data$otu_id)
print(colnames(obj_HD$data$otu_table))

obj_HD$data$otu_props <- calc_obs_props(obj_HD, "otu_table", cols = obj_HD$data$sample_data$sample_id)

obj_HD$data$tax_abund <- calc_taxon_abund(obj_HD, "otu_props", cols = obj_HD$data$sample_data$sample_id)

obj_HD$data$tax_mean_prop <- calc_group_mean(obj_HD, data = "tax_abund",
                                          cols = obj_HD$data$sample_data$sample_id,
                                          groups = "mean_prop")

obj_HD %>% 
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = mean_prop,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_HD_plot_order_RA.pdf")

```

```{r}
obj_HH = parse_phyloseq(ps_f2_HH_pruned)

print(obj_HH$data$sample_data$sample_id)
print(obj_HH$data$tax_data$otu_id)
print(colnames(obj_HH$data$otu_table))

obj_HH$data$otu_props <- calc_obs_props(obj_HH, "otu_table", cols = obj_HH$data$sample_data$sample_id)

obj_HH$data$tax_abund <- calc_taxon_abund(obj_HH, "otu_props", cols = obj_HH$data$sample_data$sample_id)

obj_HH$data$tax_mean_prop <- calc_group_mean(obj_HH, data = "tax_abund",
                                          cols = obj_HH$data$sample_data$sample_id,
                                          groups = "mean_prop")

obj_HH %>% 
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = mean_prop,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_HH_plot_order_RA.pdf")

```

Hmm, it looks like the DD samples overall are more diverse. Additionally, there are less Cyanobacteria and more Firmicutes in DD samples.

We should test alpha diversity between the different groups.

```{r}
#Alpha Diversity Analyses
#Statistical Test
data_adiv<- estimate_richness(ps_f2, measures  = c("Shannon", "InvSimpson", "Observed"))
shapiro.test(data_adiv$Shannon) #Not Normal
shapiro.test(data_adiv$InvSimpson) #Not Normal
shapiro.test(data_adiv$Observed) #Not Normal

adiv_df<- as.data.frame(ps_f2@sam_data)
adiv_df$Shannon<- data_adiv$Shannon 
adiv_df$Simpson<- data_adiv$InvSimpson
adiv_df$Observed<- data_adiv$Observed

adiv_df<- as(sample_data(adiv_df), "matrix")
adiv_df<- as.data.frame(adiv_df)
kruskal.test(Shannon ~ Condition, data=adiv_df) #Not significant
kruskal.test(Observed ~ Condition, data=adiv_df) #Not significant
kruskal.test(Simpson ~ Condition, data=adiv_df) #not significant
```

```{r}
pdf("Alpha_Div_Boxplot_Condition.pdf",width=8.25, height=6)
richness<- plot_richness(ps_f2, x="Condition", measures=c("Shannon", "InvSimpson", "Observed"))
richness$layers <- richness$layers[-1]
richness + 
  geom_boxplot( alpha = 0.6,  outlier.shape=NA, color="black", aes(fill = Condition))+
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  theme(strip.background =element_rect(fill="white"))+
  theme(axis.text.x = element_text(colour="black"))+
  theme(axis.text.y = element_text(colour="black"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1.1, vjust=1.1))+
  theme(axis.text.x = element_text(size = 14))+
  theme(axis.text.y = element_text(size = 14))+
  theme(axis.title.x = element_text(size = 18))+
  theme(axis.title.y = element_text(size = 18))+
  theme(strip.text.x=element_text(size=16, face="bold"))+
  theme(legend.title = element_text(size = 16, face = "bold"))+
  theme(legend.text = element_text(size = 12)) +
  labs(x="Condition") +
  theme(
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    panel.background = element_blank(),  # Remove panel background
    axis.line = element_line(color = "black")  # Add axis lines
  ) +
  scale_fill_manual(values = palette2) 
dev.off()
```

Which Microbes Are Enriched in Healthy vs. Diseased Tissue?

```{r}
ps_f2
ps_HH_DD<- subset_samples(ps_f2, Condition == "HH" | Condition == "DD")
ps_HH_DD

ps_HH_DD_pruned <- prune_taxa(taxa_sums(ps_HH_DD) > 0, ps_HH_DD)
ps_HH_DD_pruned

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_HH_DD_pruned)$Condition<-factor(sample_data(ps_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results.txt",sep="\t",col.names=NA)
sig_dif_hh_dd<- read.table("DESeq2_results.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many phyla
phyla<-unique(sig_dif$Phylum)
print(phyla)

tenacibaculum<- sig_dif[sig_dif$Genus =="Tenacibaculum",]
print(tenacibaculum$ASV)

vibrio<- sig_dif[sig_dif$Genus =="Vibrionaceae",]
print(vibrio$ASV)

cols <- c("Proteobacteria"="#003049", "Bdellovibrionota"="#d62828", "Bacteroidota" = "#f77f00", "Firmicutes" = "#fcbf49", "Verrucomicrobiota" = "#eae2b7")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa.pdf",width=11,height=8)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Genus,fill=Phylum))+
  geom_point(size=5, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Lowest Taxonomic Assignment")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

We're going to try this again, but colored by Order instead.

```{r}
#Count how many orders
orders<-unique(sig_dif$Order)
print(orders)

cols <- c("Rhodobacterales"="#003049", "Oligoflexales"="#d62828", "Flavobacteriales" = "#f77f00", "Mycoplasmatales" = "#fcbf49", "Verrucomicrobiales" = "#eae2b7", "Enterobacterales" = "#c8553d", "Micavibrionales" = "#588b8b")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa.pdf",width=11,height=8)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Genus,fill=Order))+
  geom_point(size=5, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Lowest Taxonomic Assignment")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

It seems like there are some patterns that may not be captured by looking at the ASV level. I will now agglomerate by different taxonomic levels to tease out some of these patterns.

```{r}
###Genus level agglomeration
ps
ps_genus<- tax_glom(ps, taxrank = "Genus")
ps_genus

#Need to Glom the taxa
ps_genus_HH_DD<- subset_samples(ps_genus, Condition == "HH" | Condition == "DD")
ps_genus_HH_DD

ps_genus_HH_DD_pruned <- prune_taxa(taxa_sums(ps_genus_HH_DD) > 0, ps_genus_HH_DD)
ps_genus_HH_DD_pruned

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_genus_HH_DD_pruned)$Condition<-factor(sample_data(ps_genus_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_genus_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_genus_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_Genus.txt",sep="\t",col.names=NA)
sig_dif<- read.table("DESeq2_results_Genus.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many phyla
cols <- c("Proteobacteria"="#003049", "Bdellovibrionota"="#d62828", "Campylobacterota"= "#6a040f","Bacteroidota" = "#f77f00", "Firmicutes" = "#fcbf49", "Verrucomicrobiota" = "#eae2b7", "Patescibacteria" = "#0a9396")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa_Genus.pdf",width=8,height=6)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Genus,fill=Phylum))+
  geom_point(size=7, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Lowest Taxonomic Assignment")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

```{r}
###Family level agglomeration
ps
ps_fam<- tax_glom(ps, taxrank = "Family")
ps_fam

#Need to Glom the taxa
ps_fam_HH_DD<- subset_samples(ps_fam, Condition == "HH" | Condition == "DD")
ps_fam_HH_DD

ps_fam_HH_DD_pruned <- prune_taxa(taxa_sums(ps_fam_HH_DD) > 0, ps_fam_HH_DD)
ps_fam_HH_DD_pruned #448 Microbial Families

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_fam_HH_DD_pruned)$Condition<-factor(sample_data(ps_fam_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_fam_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_fam_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_Family.txt",sep="\t",col.names=NA)
sig_dif<- read.table("DESeq2_results_Family.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many phyla
print(unique(sig_dif$Phylum))

cols <- c("Proteobacteria"="#003049", "Bdellovibrionota"="#d62828", "Campylobacterota"= "#6a040f","Bacteroidota" = "#f77f00", "Firmicutes" = "#fcbf49", "Verrucomicrobiota" = "#eae2b7", "Patescibacteria" = "#0a9396")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa_Family.pdf",width=8,height=6)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Family,fill=Phylum))+
  geom_point(size=7, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Family")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

```{r}
###Order level agglomeration
ps
ps_order<- tax_glom(ps, taxrank = "Order")
ps_order

#Need to Glom the taxa
ps_order_HH_DD<- subset_samples(ps_order, Condition == "HH" | Condition == "DD")
ps_order_HH_DD

ps_order_HH_DD_pruned <- prune_taxa(taxa_sums(ps_order_HH_DD) > 0, ps_order_HH_DD)
ps_order_HH_DD_pruned #256 Orders

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_order_HH_DD_pruned)$Condition<-factor(sample_data(ps_order_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_order_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_order_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_Order.txt",sep="\t",col.names=NA)
sig_dif<- read.table("DESeq2_results_Order.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many phyla
print(unique(sig_dif$Phylum))

cols <- c("Proteobacteria"="#003049", "Bdellovibrionota"="#d62828", "Bacteroidota" = "#f77f00", "Firmicutes" = "#fcbf49", "Verrucomicrobiota" = "#eae2b7", "Patescibacteria" = "#0a9396")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa_Order.pdf",width=8,height=6)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Order,fill=Phylum))+
  geom_point(size=7, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Taxonomic Order")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

Now, we Will Create a Custom Database to Match the Sequences In This Project To Those From Becker et al., 2020 and Meyer et al., 2019.

```{r}
#Read in Becker Bioindicator FASTA
bioindics<- read.fasta(file = "Bioindicator_ASVs_Becker_etal_2020.fasta", seqtype="DNA",
                       as.string = TRUE, forceDNAtolower = FALSE)
bioindics1<- unlist(bioindics)

sctld_seqs<- read.fasta(file = "5ASvs_SCTLD.fasta",seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
sctld_seqs1<- unlist(sctld_seqs)

db_seqs<- c(bioindics1, sctld_seqs)
write.fasta(as.list(db_seqs), names = names(db_seqs), file.out = "db_seqs.fasta", open = "w")


#Create FASTA out of ASVs from FLP study
#Take the rownames of the OTU Table and turn them into a FASTA file
otu<- as(otu_table(ps_f2), "matrix")
sequences<- colnames(otu)
names(sequences)<- paste('ASV',1:1202, sep="")
write.fasta(as.list(sequences), names = names(sequences), file.out = "FLP_Project_ASVs.fasta", open = "w")

blast_out<- system2(command = "/usr/local/ncbi/blast/bin/blastn", 
        args = c("-db SCTLD_db -query FLP_Project_ASVs.fasta -outfmt 6 -evalue 10e-6 -ungapped"),
        wait = TRUE,
        stdout = TRUE)

colnames <- c("qseqid",
              "sseqid",
              "pident",
              "length",
              "mismatch",
              "gapopen",
              "qstart",
              "qend",
              "sstart",
              "send",
              "evalue",
              "bitscore")

tidy_blast <- blast_out %>%
  as_tibble %>% 
  separate(col = value, 
           into = colnames,
           sep = "\t",
           convert = TRUE)

#Get the ASVs that had 100% identity
tidy_blast_100<- tidy_blast[tidy_blast$pident==100,]

#Get the Sequences that are 100% across the entire sequence length
tidy_blast_100<- tidy_blast_100[tidy_blast_100$length>=126,]

#Are any of these the same ASVs as the Enriched in Diseased Lesions When Compared to Healthy Tissue on Healthy Colonies?
#Get sequences for ASVs with 100% identity
iden_asvs<- tidy_blast_100$qseqid
iden_seqs<- sequences[names(sequences) %in% iden_asvs]
tidy_blast_100_no_repeats<- tidy_blast_100[-15,]

tidy_blast_100_no_repeats$sequence<- sequences[names(sequences) %in% tidy_blast_100$qseqid]

write.csv(tidy_blast_100_no_repeats, "blast_hits_100.csv")

sequence_blast_hits<- tidy_blast_100_no_repeats$sequence

sig_dif_hh_dd_asvs<- sig_dif_hh_dd$ASV

print(intersect(iden_seqs, sig_dif_hh_dd_asvs))

common_asvs<- intersect(iden_seqs, sig_dif_hh_dd_asvs)

#Can we get Taxonomy for these matches?

sequence_matches<- sequences[names(sequences) %in% tidy_blast_100$qseqid]

asvs_of_matches<- c()

tax<- as.data.frame(ps_f2@tax_table)
tax_of_matches<- tax[rownames(tax) %in% sequence_matches,]



```

```{r}
#Make A Relative Abundance Plot for Differentially ABundant ASVs
ps_f2

# Remove specific samples
samples_to_remove <- c("DEPBlank3-14-24", "DEPBlank3-6-24")
ps_f2_pruned <- prune_samples(!(sample_names(ps_f2) %in% samples_to_remove), ps_f2)

ps_f2_pruned

ps_f2_ra <- transform_sample_counts(ps_f2_pruned, function(OTU) OTU/sum(OTU))

dd_enriched <- subset_taxa(ps_f2_ra, row.names(ps_f2_ra@tax_table) %in% sig_dif_hh_dd_asvs)

dd_enriched

dd_melt <- psmelt(dd_enriched)

palette2<- c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")

new_labels <- setNames(dd_melt$Genus, dd_melt$OTU)

unique_asvs<- unique(dd_melt$OTU)

ggplot(dd_melt, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot() +
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black"))

```

```{r}
#Subset by ASV
dd_melt_1<- dd_melt[dd_melt$OTU == unique_asvs[1],]


dd_plot_1<- ggplot(dd_melt_1, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_1
```

```{r}
dd_melt_2<- dd_melt[dd_melt$OTU == unique_asvs[2],]

dd_plot_2<- ggplot(dd_melt_2, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_2
```

```{r}
dd_melt_3<- dd_melt[dd_melt$OTU == unique_asvs[3],]

dd_plot_3<- ggplot(dd_melt_3, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_3
```

```{r}
dd_melt_4<- dd_melt[dd_melt$OTU == unique_asvs[4],]

dd_plot_4<- ggplot(dd_melt_4, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_4

```

```{r}
dd_melt_5<- dd_melt[dd_melt$OTU == unique_asvs[5],]

dd_plot_5<- ggplot(dd_melt_5, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_5

```

```{r}
dd_melt_6<- dd_melt[dd_melt$OTU == unique_asvs[6],]

dd_plot_6<- ggplot(dd_melt_6, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_6
```

```{r}
dd_melt_7<- dd_melt[dd_melt$OTU == unique_asvs[7],]

dd_plot_7<- ggplot(dd_melt_7, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_7
```

```{r}
dd_melt_8<- dd_melt[dd_melt$OTU == unique_asvs[8],]

dd_plot_8<- ggplot(dd_melt_8, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_8
```

```{r}
dd_melt_9<- dd_melt[dd_melt$OTU == unique_asvs[9],]

dd_plot_9<- ggplot(dd_melt_9, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_9
```

```{r}
dd_melt_10<- dd_melt[dd_melt$OTU == unique_asvs[10],]

dd_plot_10<- ggplot(dd_melt_10, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Condition", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_10
```

```{r}
dd_melt_11<- dd_melt[dd_melt$OTU == unique_asvs[11],]

dd_plot_11<- ggplot(dd_melt_11, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Condition", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_11
```

```{r}
dd_melt_12<- dd_melt[dd_melt$OTU == unique_asvs[12],]

dd_plot_12<- ggplot(dd_melt_12, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Condition", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_12
```

```{r}
dd_melt_13<- dd_melt[dd_melt$OTU == unique_asvs[13],]

dd_plot_13<- ggplot(dd_melt_13, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Condition", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_13
```

```{r}
dd_melt_14<- dd_melt[dd_melt$OTU == unique_asvs[14],]

dd_plot_14<- ggplot(dd_melt_14, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Condition", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank())+
  theme(axis.title.x = element_blank())

dd_plot_14
```

```{r}
library(cowplot)

combined_plot<- plot_grid(dd_plot_1, dd_plot_2, dd_plot_3, dd_plot_4, dd_plot_5, dd_plot_6, dd_plot_7, dd_plot_8, dd_plot_9, dd_plot_10, dd_plot_11, dd_plot_12, dd_plot_13, dd_plot_14, ncol=4, nrow = 4, align="v") +
  theme(plot.margin = margin(b = 15, l=15, unit = "pt"))

combined_plot

final_plot <- ggdraw(combined_plot) +
  draw_label("Health Condition", x = 0.5, y=0.0, size = 14, vjust = -0.5) +
  draw_label("Relative Abundance", x = 0, y = 0.5, angle = 90, size = 14, vjust = 1.5)

final_plot

pdf("combined_relative_abundance_plots.pdf", width=8, height=8)

final_plot

dev.off()
```

```{r}
avg_ra<- dd_melt %>% 
  group_by(OTU, Condition) %>%
  summarise(MeanRelativeAbundance = mean(Abundance) * 100)

avg_ra_1<- avg_ra[avg_ra$OTU == unique_asvs[1],]
avg_ra_2<- avg_ra[avg_ra$OTU == unique_asvs[2],]
avg_ra_3<- avg_ra[avg_ra$OTU == unique_asvs[3],]
avg_ra_4<- avg_ra[avg_ra$OTU == unique_asvs[4],]
avg_ra_5<- avg_ra[avg_ra$OTU == unique_asvs[5],]
avg_ra_6<- avg_ra[avg_ra$OTU == unique_asvs[6],]
avg_ra_7<- avg_ra[avg_ra$OTU == unique_asvs[7],]
avg_ra_8<- avg_ra[avg_ra$OTU == unique_asvs[8],]
avg_ra_9<- avg_ra[avg_ra$OTU == unique_asvs[9],]
avg_ra_10<- avg_ra[avg_ra$OTU == unique_asvs[10],]
avg_ra_11<- avg_ra[avg_ra$OTU == unique_asvs[11],]
avg_ra_12<- avg_ra[avg_ra$OTU == unique_asvs[12],]
avg_ra_13<- avg_ra[avg_ra$OTU == unique_asvs[13],]
avg_ra_14<- avg_ra[avg_ra$OTU == unique_asvs[14],]
avg_ra_15<- avg_ra[avg_ra$OTU == unique_asvs[15],]
avg_ra_16<- avg_ra[avg_ra$OTU == unique_asvs[16],]
avg_ra_17<- avg_ra[avg_ra$OTU == unique_asvs[17],]
avg_ra_18<- avg_ra[avg_ra$OTU == unique_asvs[18],]
avg_ra_19<- avg_ra[avg_ra$OTU == unique_asvs[19],]

asv_pos_neg<- dd_melt %>% 
  group_by(OTU, Condition) %>%
  summarise(Positive = sum(Abundance > 0),
    Negative = sum(Abundance == 0),
    Total = n())

asv_pos_neg_1<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[1],]
asv_pos_neg_2<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[2],]
asv_pos_neg_3<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[3],]
asv_pos_neg_4<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[4],]
asv_pos_neg_5<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[5],]
asv_pos_neg_6<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[6],]
asv_pos_neg_7<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[7],]
asv_pos_neg_8<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[8],]
asv_pos_neg_9<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[9],]
asv_pos_neg_10<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[10],]
asv_pos_neg_11<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[11],]
asv_pos_neg_12<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[12],]
asv_pos_neg_13<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[13],]
asv_pos_neg_14<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[14],]
asv_pos_neg_15<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[15],]
asv_pos_neg_16<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[16],]
asv_pos_neg_17<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[17],]
asv_pos_neg_18<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[18],]
asv_pos_neg_19<- asv_pos_neg[asv_pos_neg$OTU == unique_asvs[19],]

```

Now we will run an indicator species analysis

```{r}
library(indicspecies)
set.seed(123)

otu <- data.frame(otu_table(ps_f2_pruned))

indic<- multipatt(otu, ps_f2_pruned@sam_data$Condition, control=how(nperm=999))
summary(indic)

DD_indicators<- indic$sign[indic$sign$s.DD == 1 & indic$sign$s.HD == 0 & indic$sign$s.HH == 0 & indic$sign$p.value < 0.05,]

DD_indicators_sorted <- DD_indicators[order(DD_indicators$stat, decreasing = TRUE), ]

DD_ind_asvs<- rownames(DD_indicators_sorted)

dd_ind_asvs_w_tax<- ps_f2_pruned@tax_table[rownames(ps_f2_pruned@tax_table) %in% DD_ind_asvs,]

dd_ind_asvs_w_tax<- as.data.frame(dd_ind_asvs_w_tax)
dd_ind_asvs_w_tax_stats<- merge(DD_indicators_sorted, dd_ind_asvs_w_tax, by=0)
dd_ind_asvs_w_tax_stats<- dd_ind_asvs_w_tax_stats[order(dd_ind_asvs_w_tax_stats$stat, decreasing = TRUE), ]
rownames(dd_ind_asvs_w_tax_stats) <- dd_ind_asvs_w_tax_stats[, 1]  
dd_ind_asvs_w_tax_stats <- dd_ind_asvs_w_tax_stats[, -1]  

matching_names <- names(sequences)[sequences %in% rownames(dd_ind_asvs_w_tax_stats)]
print(matching_names)
dd_ind_asvs_w_tax_stats$ASV<- matching_names

write.csv(dd_ind_asvs_w_tax, "diseased_indicator_species.csv")
write.csv(dd_ind_asvs_w_tax_stats, "diseased_indicator_species_sorted.csv")

top_5_dd_indics<- rownames(DD_indicators_sorted)[1:5]
top_5_dd_indics_w_tax<- dd_ind_asvs_w_tax[rownames(dd_ind_asvs_w_tax) %in% top_5_dd_indics,]


print(rownames(top_5_dd_indics_w_tax)[1])



print(rownames(top_5_dd_indics_w_tax)[5])
```

We Now Need to Rarefy the Data to Account for Uneven Sampling Depth

```{r}
rarefied_ps <- rarefy_even_depth(ps_no_contam, sample.size = 1000)

#Summarize rarefied phyloseq
summarize_phyloseq(rarefied_ps)
rarefied_ps
```

Beta Diversity with Rarefied Data

```{r}
#Beta Diversity Analyses
#Write out filtered Dataset Table
otu_ps_rare = as(otu_table(rarefied_ps), "matrix")
taxon_ps_rare = as(tax_table(rarefied_ps), "matrix")
samples_ps_rare<- as(sample_data(rarefied_ps), "matrix")
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu_ps_rare), method="CZM", label=0, z.warning = 0.999)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)

palette2<- c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")

metadata_ps_rare<- as.data.frame(samples_ps_rare)

p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=metadata_ps_rare$Condition, shape=metadata_ps_rare$Site))
p<-p+geom_point( size=3.5)+
  geom_text(aes(label = rownames(metadata_ps_rare)), vjust = 1.5, hjust = 0.5) +
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(legend.position = "right")+
  theme(legend.title = element_blank()) +
  theme(legend.key.width = unit(1.0,  unit = "cm")) +
  guides(fill = guide_legend(byrow = TRUE)) 
p<- p+ labs(x=xlab, y=ylab, fill="Condition") + coord_fixed()+
  labs(fill = "Condition", shape="Site")+
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_fill_manual(values = palette2) +
  theme_minimal() +
  labs(x = "Principal Component 1: 30%", y = "Principal Component 2: 5%") +
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))
p
```

Just to explore without the one outlier.....

```{r}
#Remove one outlier sample
samples_to_remove<- "FLP73"

rarefied_ps_no_73<- prune_samples(!(sample_names(rarefied_ps) %in% samples_to_remove), rarefied_ps)

rarefied_ps_no_73

#Beta Diversity Analyses
#Write out filtered Dataset Table
otu_ps_rare = as(otu_table(rarefied_ps_no_73), "matrix")
taxon_ps_rare = as(tax_table(rarefied_ps_no_73), "matrix")
samples_ps_rare<- as(sample_data(rarefied_ps_no_73), "matrix")
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu_ps_rare), method="CZM", label=0, z.warning = 0.999)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
df_out <- as.data.frame(d.pcx$x)

palette2<- c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")

metadata_ps_rare<- as.data.frame(samples_ps_rare)

p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=metadata_ps_rare$Condition, shape=metadata_ps_rare$Site))
p<-p+geom_point( size=3.5)+
  geom_text(aes(label = rownames(metadata_ps_rare)), vjust = 1.5, hjust = 0.5) +
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(legend.position = "right")+
  theme(legend.title = element_blank()) +
  theme(legend.key.width = unit(1.0,  unit = "cm")) +
  guides(fill = guide_legend(byrow = TRUE)) 
p<- p+ labs(x=xlab, y=ylab, fill="Condition") + coord_fixed()+
  labs(fill = "Condition", shape="Site")+
  scale_shape_manual(values = c(21, 22, 23, 24, 25)) +
  scale_fill_manual(values = palette2) +
  theme_minimal() +
  labs(x = "Principal Component 1: 30%", y = "Principal Component 2: 5%") +
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))
p
```

```{r}
otu_ps_rare = as(otu_table(rarefied_ps), "matrix")
taxon_ps_rare = as(tax_table(rarefied_ps), "matrix")
samples_ps_rare<- as(sample_data(rarefied_ps), "matrix")

d.czm <- cmultRepl(t(otu_ps_rare), method="CZM", label=0, z.warning = 0.999)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
summary(d.pcx)
str(d.pcx)
screeplot(d.pcx)

#Run the PERMANOVA w Site and Condition
###################PERMANOVA#################
dist.clr <- dist(E.clr)
perm<-adonis2(dist.clr~Condition+Site,as(sample_data(rarefied_ps),"data.frame"))
print(perm)

write.table(perm, "permanova_results_rarefied.txt")

### Beta Diversity Dispersions - Condition
#calculate multivariate dispersions based on Condition
metadata_ps_rare<- as.data.frame(samples_ps_rare)
mod <-betadisper(dist.clr, metadata_ps_rare$Condition)
#one way anova
anova(mod)
#boxplots
pdf("beta_dispersion_condition_rarefied.pdf",width=8.5)
plot(mod)
dev.off()
```

```{r}
## Compute mean distance to centroid per group
#this just prints values on the console
tapply(mod$distances, metadata_ps_rare$Condition, mean)
## Same, but variance instead
tapply(mod$distances, metadata_ps_rare$Condition, var)

#Get the distances to centroid from the model
mod$distances
dis <- mod$distances
#melt
dis.melt <- melt(dis)
#move rownames to columns so we can merge the dispersion values and metadata
dis.melt$Sample <- rownames(dis.melt)
samples$Sample <- rownames(samples)
#merge metadata and dispersion 
dis.condition <- merge(samples, dis.melt)
#rename column
colnames(dis.condition)[4] <- "distance"

shapiro.test(dis.condition$distance) #Not normally distributed, use kruskal test
distance.test<- kruskal.test(distance ~ Condition, data=dis.condition)
print(distance.test)

capture.output(distance.test,file="kruskal_distance_to_centroid_rarefied.txt")

#Post Hoc Test
library(FSA)
dunnTest(distance ~ Condition,
  data = dis.condition,
  method = "bonferroni"
)
```

```{r}
#Create Violin plot for distance to centroid for rarefied data
library(ggpubr)
p2<-ggplot(dis.condition, aes(x=Condition, y=distance)) +
  geom_violin(mapping =aes(x=Condition,y=distance, fill=Condition, color="black"), alpha = 0.6) +
  geom_point(position=position_jitter(width=.1, height=0),aes(colour="black",fill=Condition),size=4, pch=21, colour = "black") +
  scale_fill_manual(values = palette2) +
  scale_color_manual(values = c( "black")) +
  ylab("Distance to Centroid")+
  xlab("Health Condition")+
  guides(color="none") +
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(colour="black"))+
  theme(axis.text.y = element_text(colour="black"))+
  theme(axis.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))+
  theme(axis.title.x = element_text(size = 16))+
  theme(axis.title.y = element_text(size = 16))+
  theme(strip.text.x=element_text(size=16, face="bold"))+
  theme(legend.position = "none") +
  stat_compare_means(
    comparisons = list(c("DD", "HH")),
    method = "t.test",
    label = "p.signif")
p2
```

Now to redo the alpha diversity analyses....

```{r}
#Alpha Diversity Analyses
#Statistical Test
data_adiv<- estimate_richness(rarefied_ps, measures  = c("Shannon", "InvSimpson", "Observed"))
shapiro.test(data_adiv$Shannon) #Not Normal
shapiro.test(data_adiv$InvSimpson) #Not Normal
shapiro.test(data_adiv$Observed) #Not Normal

adiv_df<- as.data.frame(rarefied_ps@sam_data)
adiv_df$Shannon<- data_adiv$Shannon 
adiv_df$Simpson<- data_adiv$InvSimpson
adiv_df$Observed<- data_adiv$Observed

adiv_df<- as(sample_data(adiv_df), "matrix")
adiv_df<- as.data.frame(adiv_df)
kruskal.test(Shannon ~ Condition, data=adiv_df) #Not significant
kruskal.test(Observed ~ Condition, data=adiv_df) #Not significant
kruskal.test(Simpson ~ Condition, data=adiv_df) #not significant
```

```{r}
pdf("Alpha_Div_Boxplot_Condition_Rarefied.pdf",width=8.25, height=6)
richness<- plot_richness(rarefied_ps, x="Condition", measures=c("Shannon", "InvSimpson", "Observed"))
richness$layers <- richness$layers[-1]
richness + 
  geom_boxplot( alpha = 0.6,  outlier.shape=NA, color="black", aes(fill = Condition))+
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  theme(strip.background =element_rect(fill="white"))+
  theme(axis.text.x = element_text(colour="black"))+
  theme(axis.text.y = element_text(colour="black"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1.1, vjust=1.1))+
  theme(axis.text.x = element_text(size = 14))+
  theme(axis.text.y = element_text(size = 14))+
  theme(axis.title.x = element_text(size = 18))+
  theme(axis.title.y = element_text(size = 18))+
  theme(strip.text.x=element_text(size=16, face="bold"))+
  theme(legend.title = element_text(size = 16, face = "bold"))+
  theme(legend.text = element_text(size = 12)) +
  labs(x="Condition") +
  theme(
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    panel.background = element_blank(),  # Remove panel background
    axis.line = element_line(color = "black")  # Add axis lines
  ) +
  scale_fill_manual(values = palette2) 
dev.off()
```

```{r}
#Create Metacoder Plot
library(metacoder)
obj_rare = parse_phyloseq(rarefied_ps)

obj_rare$data$otu_props <- calc_obs_props(obj_rare, "otu_table", cols = obj_rare$data$sample_data$sample_id)

obj_rare$data$tax_abund <- calc_taxon_abund(obj_rare, "otu_props", cols = obj_rare$data$sample_data$sample_id)

obj_rare$data$tax_mean_prop <- calc_group_mean(obj_rare, data = "tax_abund",
                                          cols = obj_rare$data$sample_data$sample_id,
                                          groups = "mean_prop")

obj_rare %>% 
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>% # subset to the order rank
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = mean_prop,
            initial_layout = "re", layout = "da",
            output_file = "metacoder_plot_order_RA_Rarefied.pdf")
```

```{r}
#Try to create a grid faceted by health condition
library(gridExtra)
obj_rare
groups <- unique(sample_data(rarefied_ps)$Condition)

my_plots <- lapply(c("HH", "HD", "DD"), function(g) {
  # Create heat tree for each time point
  set.seed(1) # For consistent layout
  
  x_subset <- filter_obs(obj_rare, obj_rare$data$sample_data$Condition==g)
  
  x_subset %>%
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
     filter_obs(sample_data, Condition == "HH") %>%
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = mean_prop,
            initial_layout = "re", layout = "da")
})

my_plots


```

```{r}
#Is Vallitalea present?
ps_f2_ra_valli<-subset_taxa(ps_f2_ra, Genus == "Vallitalea")
ps_f2_ra_valli

otu_table_valli <- as.data.frame(otu_table(ps_f2_ra_valli))
samples_valli<- as.data.frame(sample_data(ps_f2_ra_valli))

plot_data <- otu_table_valli %>%
  rownames_to_column("Sample") %>%
  pivot_longer(cols = -Sample, names_to = "ASV", values_to = "Abundance")

plot_metadata<- cbind(Sample = rownames(samples_valli), samples_valli)
rownames(plot_metadata) <- NULL

valli_plot<- merge(plot_data, plot_metadata, by ="Sample")

new_labels <- c("TACGTAGGGGGCAAGCGTTATCCGGATTTACTGGGTGTAAAGGGTGCGTAGGCGGCGAAGTAAGTCAGATGTGAAAGCCCGAGGCTCAACCTCGGGACTGCATTTGAAACTGCTTTGCTAGAGTGCAGGAGAGGAAAGCGGAATTCCTAGTGTAGCGGTGAAATGCGTAGATATTAGGAGGAACACCAGTGGCGAAGGCGGCTTTCTGGACTGTAACTGACGCTGAGGCACGAAAGCGTGGGGAGCGAACAGG" = "Vallitalea ASV1", "TACGTAGGGGGCAAGCGTTATCCGGATTTACTGGGTGTAAAGGGTGCGTAGGCGGCTAGGTAAGTCAGATGTGAAAGCCCGAGGCTCAACCTCGGGACTGCATTTGAAACTGCTTAGCTAGAGTGCAGGAGAGGAAAGTGGAATTCCTAGTGTAGCGGTGAAATGCGTAGATATTAGGAGGAACACCAGTGGCGAAGGCGACTTTCTGGACTGTAACTGACGCTGAGGCACGAAAGCGTGGGGAGCGAACAGG" = "Vallitalea ASV 2")

ggplot(valli_plot, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(shape = 21, width = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Health Condition", y = "Relative Abundance", fill = "Condition") +
  scale_y_continuous(labels = scales::percent) +
  facet_grid(~ASV, labeller = labeller(ASV = new_labels)) + 
  scale_fill_manual(values = palette2)

```

```{r}
#Is Wukongibacter present?
ps_f2_ra
ps_f2_ra_wuko<-subset_taxa(ps_f2_ra, Genus == "Wukongibacter")
#No Wukongibacter present in dataset

```

```{r}
#What about Clostridium sensu stricto 1?
ps_f2_ra_clost<-subset_taxa(ps_f2_ra, Genus == "Clostridium sensu stricto 1")
ps_f2_ra_clost

rownames(ps_f2_ra_clost@tax_table)

otu_table_valli <- as.data.frame(otu_table(ps_f2_ra_valli))
samples_valli<- as.data.frame(sample_data(ps_f2_ra_valli))
```

```{r}
length(unique(rownames(samples)))
```

Given that there are certain indicators of SCTLD in Unaffected tissue from diseased colonies, it should be worth doing another analysis between HD and HH samples, as well as HD and DD.

```{r}
ps_f2

#Need to subset samples by the health conditions
ps_f2_HH_HD<- subset_samples(ps_f2, Condition == "HH" | Condition == "HD")
ps_f2_HH_HD

ps_f2_HH_HD_pruned <- prune_taxa(taxa_sums(ps_f2_HH_HD) > 0, ps_f2_HH_HD)
ps_f2_HH_HD_pruned #1119 ASVs

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Unaffected Tissue
sample_data(ps_f2_HH_HD_pruned)$Condition<-factor(sample_data(ps_f2_HH_HD_pruned)$Condition,levels=c("HH","HD"))
head(sample_data(ps_f2_HH_HD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_f2_HH_HD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
```

THere are no significantly different ASVs between HH and HD using the filtered phyloseq object. How about the unfiltered phyloseq object?

```{r}
ps

#Need to subset samples by the health conditions
ps_HH_HD<- subset_samples(ps, Condition == "HH" | Condition == "HD")
ps_HH_HD

ps_HH_HD_pruned <- prune_taxa(taxa_sums(ps_HH_HD) > 0, ps_HH_HD)
ps_HH_HD_pruned #3941 ASVs

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Unaffected Tissue
sample_data(ps_HH_HD_pruned)$Condition<-factor(sample_data(ps_HH_HD_pruned)$Condition,levels=c("HH","HD"))
head(sample_data(ps_HH_HD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_HH_HD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)

#Still no significantly different ASVs when using unfiltered dataset
#This isn't unexpected, given the HH and HD samples were not significantly different from each other in the beta diversity dispersions
```

#Let's try it again but looking at the HD and DD samples

```{r}
ps

#Need to subset samples by the health conditions
ps_DD_HD<- subset_samples(ps, Condition == "DD" | Condition == "HD")
ps_DD_HD

ps_DD_HD_pruned <- prune_taxa(taxa_sums(ps_DD_HD) > 0, ps_DD_HD)
ps_DD_HD_pruned #4195 ASVs

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_DD_HD_pruned)$Condition<-factor(sample_data(ps_DD_HD_pruned)$Condition,levels=c("HD","DD"))
head(sample_data(ps_DD_HD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_DD_HD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)

#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_ASV_HD_DD.txt",sep="\t",col.names=NA)
sig_dif<- read.table("DESeq2_results_ASV_HD_DD.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many phyla
print(unique(sig_dif$Phylum))

cols <- c("Proteobacteria"="#003049", "Bdellovibrionota"="#d62828", "Bacteroidota" = "#f77f00", "Firmicutes" = "#fcbf49", "Verrucomicrobiota" = "#eae2b7", "Campylobacterota" = "#0a9396")

flavo<- sig_dif[sig_dif$Family=="Flavobacteriaceae",]
flavo$ASV

rickett<- sig_dif[sig_dif$Family=="Rickettsiales",]
rickett$ASV

vibrio<- sig_dif[sig_dif$Genus=="Vibrio",]
vibrio$ASV

#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa_ASV_HD_DD.pdf",width=8,height=6)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Genus,fill=Phylum))+
  geom_point(size=7, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Taxonomic Order")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

Are any of these present in the blast hits?

```{r}
sequence_matches<- sig_dif[sig_dif$ASV %in% sequence_blast_hits,]
sequence_matches<- tidy_blast_100_no_repeats[tidy_blast_100_no_repeats$sequence %in% sig_dif$ASV,]
```

Now to include more of the differentially abundant ASVs between diseased and unaffected tissue

```{r}
#Subset by ASV
samples_to_remove <- c("DEPBlank3-14-24", "DEPBlank3-6-24")
ps_f2_pruned <- prune_samples(!(sample_names(ps_f2) %in% samples_to_remove), ps_f2)

ps_f2_pruned

ps_f2_ra <- transform_sample_counts(ps_f2_pruned, function(OTU) OTU/sum(OTU))

dd_enriched_2 <- subset_taxa(ps_f2_ra, row.names(ps_f2_ra@tax_table) %in% sequence_matches$sequence)

dd_melt_2 <- psmelt(dd_enriched_2)

dd_melt_halarco<- dd_melt[dd_melt$OTU == sequence_matches$sequence[1],]


ggplot(dd_melt_halarco, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) 

avg_ra<- dd_melt_2 %>% 
  group_by(OTU, Condition) %>%
  summarise(MeanRelativeAbundance = mean(Abundance) * 100)

avg_ra_1<- avg_ra[avg_ra$OTU == sequence_matches$sequence[1],]

asv_pos_neg<- dd_melt_2 %>% 
  group_by(OTU, Condition) %>%
  summarise(Positive = sum(Abundance > 0),
    Negative = sum(Abundance == 0),
    Total = n())

asv_pos_neg_1<- asv_pos_neg[asv_pos_neg$OTU == sequence_matches$sequence[1],]

```

```{r}
dd_melt_plankto<- dd_melt[dd_melt$OTU == sequence_matches$sequence[3],]


ggplot(dd_melt_plankto, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) 

avg_ra_3<- avg_ra[avg_ra$OTU == sequence_matches$sequence[3],]

asv_pos_neg_3<- asv_pos_neg[asv_pos_neg$OTU == sequence_matches$sequence[3],]
```

What's Highest Relative Abundance in Healthy Tissue?

```{r}
#ASV Level

ps_f2_ra

ps_f2_ra_hh<- subset_samples(ps_f2_ra, Condition=="HH")

hh_melt <- psmelt(ps_f2_ra_hh)

avg_ra_hh<- hh_melt %>% 
  group_by(OTU, Condition) %>%
  summarise(MeanRelativeAbundance = mean(Abundance) * 100)

avg_ra_hh_w_tax<- merge(avg_ra_hh, hh_melt, by = "OTU")

avg_ra_hh_w_tax_sorted<- avg_ra_hh_w_tax[order(avg_ra_hh_w_tax$MeanRelativeAbundance, decreasing = TRUE), ] 

avg_ra_hh_w_tax_sorted<-avg_ra_hh_w_tax_sorted[avg_ra_hh_w_tax_sorted$MeanRelativeAbundance > 0,]

length(unique(avg_ra_hh_w_tax_sorted$OTU))
```

What's highest relative abundance in Diseased Tissue?

```{r}
#ASV Level

ps_f2_ra

ps_f2_ra_dd<- subset_samples(ps_f2_ra, Condition=="DD")

dd_melt <- psmelt(ps_f2_ra_dd)

avg_ra_dd<- dd_melt %>% 
  group_by(OTU, Condition) %>%
  summarise(MeanRelativeAbundance = mean(Abundance) * 100)

avg_ra_dd_w_tax<- merge(avg_ra_dd, dd_melt, by = "OTU")

avg_ra_dd_w_tax_sorted<- avg_ra_dd_w_tax[order(avg_ra_dd_w_tax$MeanRelativeAbundance, decreasing = TRUE), ] 

avg_ra_dd_w_tax_sorted<-avg_ra_dd_w_tax_sorted[avg_ra_dd_w_tax_sorted$MeanRelativeAbundance > 0,]

length(unique(avg_ra_dd_w_tax_sorted$OTU))
```

###Let's Rerun our DESeq2 Results with the filtered dataset

```{r}
###Order level agglomeration
ps_f2
ps_f2_order<- tax_glom(ps_f2, taxrank = "Order")
ps_f2_order

#Need to Glom the taxa
ps_f2_order_HH_DD<- subset_samples(ps_f2_order, Condition == "HH" | Condition == "DD")
ps_f2_order_HH_DD

ps_f2_order_HH_DD_pruned <- prune_taxa(taxa_sums(ps_f2_order_HH_DD) > 0, ps_f2_order_HH_DD)
ps_f2_order_HH_DD_pruned #139 Orders

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_f2_order_HH_DD_pruned)$Condition<-factor(sample_data(ps_f2_order_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_f2_order_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis
dds = phyloseq_to_deseq2(ps_f2_order_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_Order.txt",sep="\t",col.names=NA)
sig_dif<- read.table("DESeq2_results_Order.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many phyla
print(unique(sig_dif$Order))

cols <- c("Rhodobacterales"="#003049", "Oligoflexales"="#d62828", "Flavobacteriales" = "#f77f00", "Mycoplasmatales" = "#fcbf49", "Verrucomicrobiales" = "#eae2b7", "Pseudomonadales" = "#287271", "Peptostreptococcales-Tissierellales" = "#ef476f", "Rickettsiales" = "#118ab2", "Gracilibacteria" = "#06d6a0", "Bacteroidales" = "#83d483")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa_Order.pdf",width=8,height=6)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Order,fill=Order))+
  geom_point(size=7, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Taxonomic Order")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank()) + 
  theme(legend.position = "none")
p1
dev.off()
```

```{r}
ps_f2
ps_f2_HH_DD<- subset_samples(ps_f2, Condition == "HH" | Condition == "DD")
ps_f2_HH_DD

ps_f2_HH_DD_pruned <- prune_taxa(taxa_sums(ps_f2_HH_DD) > 0, ps_f2_HH_DD)
ps_f2_HH_DD_pruned

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_f2_HH_DD_pruned)$Condition<-factor(sample_data(ps_f2_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_f2_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_f2_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results.txt",sep="\t",col.names=NA)
sig_dif_hh_dd<- read.table("DESeq2_results.txt", header = TRUE)
colnames(sig_dif_hh_dd)[1]<- "ASV"

#Count how many orders
orders<-unique(sig_dif_hh_dd$Order)
print(orders)

cols <- c("Rhodobacterales"="#003049", "Oligoflexales"="#d62828", "Flavobacteriales" = "#f77f00", "Mycoplasmatales" = "#fcbf49", "Verrucomicrobiales" = "#eae2b7", "Enterobacterales" = "#c8553d", "Micavibrionales" = "#588b8b", "Campylobacterales" = "#bc6c25")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa.pdf",width=11,height=8)
p1 <- ggplot(sig_dif_hh_dd,aes(log2FoldChange,Genus,fill=Order))+
  geom_point(size=5, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Lowest Taxonomic Assignment")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank()) +
  scale_x_continuous(limits = c(-5, 30)) +
  scale_y_discrete(limits= c("Vibrionaceae", "Vibrio", "Thalassobius", "Tenacibaculum", "Ruegeria", "Rubritalea","Rhodobacteraceae", "Pseudobacteriovorax", "Mycoplasma","Micavibrionaceae", "Halarcobacter", "Ferrimonas", "Cryomorphaceae", "Cribrihabitans", "Aurantivirga"),
                   labels= c("Vibrionaceae", expression(italic("Vibrio")), expression(italic("Thalassobius")), 
                             expression(italic("Tenacibaculum")), 
                             expression(italic("Ruegeria")), expression(italic("Rubritalea")), "Rhodobacteraceae",
                             expression(italic("Pseudobacteriovorax")),
                             expression(italic("Mycoplasma")),
                             "Micavibrionaceae", expression(italic("Halarcobacter")), 
                             expression(italic("Ferrimonas")), "Cryomorphaceae",
                             expression(italic("Cribrihabitans")), 
                            expression(italic("Aurantivirga"))))
p1
dev.off()
```

Let's redo the relative abundance plots to include those that are considered differentialy abundant in the filtered dataset.

### Differential Abundance with ps_f2 between DD and HD

```{r}
ps_f2
ps_f2_HD_DD<- subset_samples(ps_f2, Condition == "HD" | Condition == "DD")
ps_f2_HD_DD

ps_f2_HD_DD_pruned <- prune_taxa(taxa_sums(ps_f2_HD_DD) > 0, ps_f2_HD_DD)
ps_f2_HD_DD_pruned

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_f2_HD_DD_pruned)$Condition<-factor(sample_data(ps_f2_HD_DD_pruned)$Condition,levels=c("HD","DD"))
head(sample_data(ps_f2_HD_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_f2_HD_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_HD_DD.txt",sep="\t",col.names=NA)
sig_dif_hd_dd<- read.table("DESeq2_results_HD_DD.txt", header = TRUE)
colnames(sig_dif_hd_dd)[1]<- "ASV"

unique_rows <- sig_dif_hd_dd[!sig_dif_hd_dd$ASV %in% sig_dif_hh_dd$ASV, ]
```

```{r}
#Make A Relative Abundance Plot for Differentially ABundant ASVs
sig_dif_hh_dd_asvs<- sig_dif_hh_dd$ASV

dd_enriched <- subset_taxa(ps_f2_ra, row.names(ps_f2_ra@tax_table) %in% sig_dif_hh_dd_asvs)

dd_enriched

dd_melt <- psmelt(dd_enriched)

palette2<- c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")

new_labels <- setNames(dd_melt$Genus, dd_melt$OTU)

unique_asvs<- unique(dd_melt$OTU)

ggplot(dd_melt, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot() +
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black"))
```

```{r}
dd_melt_1<- dd_melt[dd_melt$OTU == unique_asvs[1],]


dd_plot_1<- ggplot(dd_melt_1, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_1
```

```{r}
dd_melt_2<- dd_melt[dd_melt$OTU == unique_asvs[2],]


dd_plot_2<- ggplot(dd_melt_2, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_2
```

```{r}
dd_melt_3<- dd_melt[dd_melt$OTU == unique_asvs[3],]


dd_plot_3<- ggplot(dd_melt_3, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_3
```

```{r}
dd_melt_4<- dd_melt[dd_melt$OTU == unique_asvs[4],]


dd_plot_4<- ggplot(dd_melt_4, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_4
```

```{r}
dd_melt_5<- dd_melt[dd_melt$OTU == unique_asvs[5],]


dd_plot_5<- ggplot(dd_melt_5, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_5
```

```{r}
dd_melt_6<- dd_melt[dd_melt$OTU == unique_asvs[6],]


dd_plot_6<- ggplot(dd_melt_6, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_6
```

```{r}
dd_melt_7<- dd_melt[dd_melt$OTU == unique_asvs[7],]


dd_plot_7<- ggplot(dd_melt_7, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_7
```

```{r}
dd_melt_8<- dd_melt[dd_melt$OTU == unique_asvs[8],]

custom_labels <- c("TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTCAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG" = "Cribrihabitans ASV21")


dd_plot_8<- ggplot(dd_melt_8, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_8
```

```{r}
dd_melt_9<- dd_melt[dd_melt$OTU == unique_asvs[9],]

custom_labels <- c("TACGGAGGGTGCGAGCGTTAATCGGAATCACTGGGCGTAAAGCGCATGCAGGCGGTTTGCTAAGCGAGATGTGAAAGCCCCGGGCTTAACCTGGGAACTGCATTTCGAACTGGCGAACTAGAGTCTTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCAGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTCAGATGCGAAAGCGTGGGGAGCAAACAGG" = "Ferrimonas ASV64")


dd_plot_9<- ggplot(dd_melt_9, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_9
```

```{r}
dd_melt_10<- dd_melt[dd_melt$OTU == unique_asvs[10],]


dd_plot_10<- ggplot(dd_melt_10, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_10
```

```{r}
dd_melt_11<- dd_melt[dd_melt$OTU == unique_asvs[11],]


dd_plot_11<- ggplot(dd_melt_11, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_11
```

```{r}
dd_melt_12<- dd_melt[dd_melt$OTU == unique_asvs[12],]


dd_plot_12<- ggplot(dd_melt_12, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_12
```

```{r}
dd_melt_13<- dd_melt[dd_melt$OTU == unique_asvs[13],]


dd_plot_13<- ggplot(dd_melt_13, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_13
```

```{r}
dd_melt_14<- dd_melt[dd_melt$OTU == unique_asvs[14],]

custom_labels <- c("TACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGTCAATTAAGTCAGAGGTGAAATCCCATAGCTTAACTATGGAACTGCCTTTGATACTGGTTGACTTGAGTTATACGGAAGTAGATAGAATAAGTAGTGTAGCGGTGAAATGCATAGATATTACTTAGAATACCGATTGCGAAGGCAGTCTACTACGTATATACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG" = "Aurantivirga ASV123")


dd_plot_14<- ggplot(dd_melt_14, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_14
```

```{r}
dd_melt_15<- dd_melt[dd_melt$OTU == unique_asvs[15],]

custom_labels <- c("TACATAGGCCCCAAGCGTTATCCGGATTTACTGGGCGTAAAGAGTGCGCAGGTGATTTCACAAGTCTGTATTGAAAGGCTGGAGCTCAACTCCAGTTTGGTATGGAAACTGTGTTATTAGATTATGATAGAGGTTAGCGGAATTCCTAATGTAGCGGTGGAATGCGTAGATATTAGGAGGAACACCAAAAGCGAAGGCAGCTAACTGGGTCATTAATGACACTGGTGCACGAAAGCGTGGGGAGCAAACAGG" = "Mycoplasma ASV80")


dd_plot_15<- ggplot(dd_melt_15, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_15
```

```{r}
dd_melt_16<- dd_melt[dd_melt$OTU == unique_asvs[16],]


dd_plot_16<- ggplot(dd_melt_16, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_16
```

```{r}
dd_melt_17<- dd_melt[dd_melt$OTU == unique_asvs[17],]


dd_plot_17<- ggplot(dd_melt_17, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_17
```

```{r}
dd_melt_18<- dd_melt[dd_melt$OTU == unique_asvs[18],]


dd_plot_18<- ggplot(dd_melt_18, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = new_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_18
```

```{r}
dd_melt_19<- dd_melt[dd_melt$OTU == unique_asvs[19],]

custom_labels <- c("TACTGGGGGAGCGAGCGTTGTTCGGAATCACTGGGCGTAAAGCGTGCGCAGGCGGTTTGGAAAGTCAGAAGTGAAATCCCAGAGCTCAACTCTGGAACTGCTTTTGAAACTTCCTTACTAGAGTTTGGTAGAGGTTGGGAGAATTCCTAGTGTAGAGGTGAAATTCGTAGATATTAGGAGGAATACCAGTGGCGTAGGCGCCCAACTGGACCATTACTGACGCTCAGGCACGAAAGCGTGGGGAGCAAACAGG" = "Micavibrionaceae ASV158")


dd_plot_19<- ggplot(dd_melt_19, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_19
```

```{r}
library(cowplot)

combined_plot<- plot_grid(dd_plot_1, dd_plot_2, dd_plot_3, dd_plot_4, dd_plot_5, dd_plot_6, dd_plot_7, dd_plot_8, dd_plot_9, dd_plot_10, dd_plot_11, dd_plot_12, dd_plot_13, dd_plot_14, dd_plot_15, dd_plot_16, dd_plot_17, dd_plot_18, dd_plot_19, ncol=4, nrow = 5, align="v") +
  theme(plot.margin = margin(b = 15, l=15, unit = "pt"))

combined_plot

final_plot <- ggdraw(combined_plot) +
  draw_label("Health Condition", x = 0.5, y=0.0, size = 14, vjust = -0.5) +
  draw_label("Relative Abundance", x = 0, y = 0.5, angle = 90, size = 14, vjust = 1.5)

final_plot

pdf("combined_relative_abundance_plots_updated.pdf", width=8, height=8)
final_plot
dev.off()
```

Since we redid the DESeq2 results, let's now recheck if any of these match with the Blast matches

```{r}
common_asvs<- intersect(iden_seqs, sig_dif_asvs)

#Can we get Taxonomy for these matches?

sequence_matches<- sequences[names(sequences) %in% tidy_blast_100$qseqid]

tax<- as.data.frame(ps_f2@tax_table)
tax_of_matches<- tax[rownames(tax) %in% sequence_matches,]

tax_of_common_asvs<- tax_of_matches[rownames(tax_of_matches) %in% common_asvs,]
```

Now let's figure out which ASVs were the same between the DESeq2, the BLAST, and the indicator analysis.

```{r}
asvs_shared<- tax_of_common_asvs[rownames(tax_of_common_asvs) %in% rownames(dd_ind_asvs_w_tax),]

#How many of the indicator species are considered differentially enriched by DESeq2?
length(sig_dif_asvs)
asvs_indics_enriched<- sig_dif_asvs[rownames(dd_ind_asvs_w_tax) %in% sig_dif_asvs]

#13 differentially enriched taxa were indicators of FLP

dd_melt_shared<- dd_melt[dd_melt$OTU %in% asvs_indics_enriched,]
dd_melt_shared<- dd_melt_shared[dd_melt_shared$OTU %in% common_asvs,]

length(unique(dd_melt_shared$OTU))
```

OK that got a little confusing, let's try this again

```{r}
indic_asvs<- rownames(dd_ind_asvs_w_tax)
enriched_asvs<- sig_dif_asvs
blast_asvs<- tidy_blast_100_no_repeats$sequence

shared_asvs<- intersect(intersect(indic_asvs, enriched_asvs), blast_asvs)

tax<- as.data.frame(ps_f2@tax_table)
tax_of_shared<- tax[rownames(tax) %in% shared_asvs,]
  
  # Remove specific samples
samples_to_remove <- c("DEPBlank3-14-24", "DEPBlank3-6-24")
ps_f2_pruned <- prune_samples(!(sample_names(ps_f2) %in% samples_to_remove), ps_f2)

ps_f2_pruned

ps_f2_ra <- transform_sample_counts(ps_f2_pruned, function(OTU) OTU/sum(OTU))

f2_pruned_ra_shared <- subset_taxa(ps_f2_ra, row.names(ps_f2_ra@tax_table) %in% shared_asvs)

f2_pruned_ra_shared

dd_melt_shared <- psmelt(f2_pruned_ra_shared)

palette2<- c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")

new_labels <- setNames(dd_melt_shared$Genus, dd_melt_shared$OTU)

unique_asvs<- unique(dd_melt_shared$OTU)
```

Let's only do relative abundance plots of the four ASVs.

```{r}
dd_melt_shared_1<- dd_melt_shared[dd_melt_shared$OTU == unique_asvs[1],]
print(unique_asvs[1])

custom_labels <- c("TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTAGTAAGTTAGAGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTTAATACTGCTAGTCTTGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG" = "Thalassobius ASV36")

dd_plot_shared_1<- ggplot(dd_melt_shared_1, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_shared_1
```

```{r}
dd_melt_shared_2<- dd_melt_shared[dd_melt_shared$OTU == unique_asvs[2],]

print(unique_asvs[2])

custom_labels <- c("TACGGAGGGTGCAAGCGTTACTCGGAATCACTGGGCGTAAAGAGCGTGTAGGCGGGTTAATAAGTCAGATGTGAAAGCCTATGGCTCAACCATAGAACTGCATTTGAAACTGTTAACCTAGAATATGGGAGAGGTAGATGGAATTTCTGGTGTAGGGGTAAAATCCGTAGAGATCAGAAGGAATACCGATTGCGAAGGCGATCTACTGGAACATTATTGACGCTGAGACGCGAAAGCGTGGGGAGCAAACAGG" = "Halarcobacter ASV25")


dd_plot_shared_2<- ggplot(dd_melt_shared_2, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_shared_2
```

```{r}
dd_melt_shared_3<- dd_melt_shared[dd_melt_shared$OTU == unique_asvs[3],]

print(unique_asvs[3])

custom_labels <- c("TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTAGGGGTGAAATCCCAGGGCTCAACCCTGGAACTGCCCTTAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG" = "Rhodobactereaceae ASV14")

dd_plot_shared_3<- ggplot(dd_melt_shared_3, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_shared_3
```

```{r}
dd_melt_shared_4<- dd_melt_shared[dd_melt_shared$OTU == unique_asvs[4],]

print(unique_asvs[4])

custom_labels <- c("TACGGAGGGTCCAAGCGTTATCCGGATTTATTGGGTTTAAAGGGTCCGTAGGCGGGGTTTTAAGTCAGTGGTGAAAGCCTACAGCTCAACTGTAGAACTGCCATTGAAACTGGAACTCTTGAATGTGATTGAGGTAGGCGGAATATGTCATGTAGCGGTGAAATGCTTAGATATGACATAGAACACCGATAGCGAAGGCAGCTTACCAAGTCATTATTGACGCTGATGGACGAAAGCGTGGGGAGCGAACAGG" = "Cryomorphaceae ASV23")


dd_plot_shared_4<- ggplot(dd_melt_shared_4, aes(x = Condition, y = Abundance, fill = Condition)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Condition)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Condition, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  scale_fill_manual(values = palette2) +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_shared_4
```

Let's now make the plot!

```{r}
combined_plot<- plot_grid(dd_plot_shared_1, dd_plot_shared_2, dd_plot_shared_3, dd_plot_shared_4, ncol=2, nrow = 2, align="v") +
  theme(plot.margin = margin(b = 15, l=15, unit = "pt"))

combined_plot

final_plot <- ggdraw(combined_plot) +
  draw_label("Health Condition", x = 0.5, y=0.0, size = 14, vjust = -0.5) +
  draw_label("Relative Abundance", x = 0, y = 0.5, angle = 90, size = 14, vjust = 1.5)

final_plot

pdf("combined_relative_abundance_plots_updated_only_common_asvs.pdf", width=8, height=6)
final_plot
dev.off()
```

Let's redo the metacoder trees, but maybe we can filter out low abundance taxa?

```{r}
#Healthy HH samples
obj_HH = parse_phyloseq(ps_f2_HH_pruned)

print(obj_HH$data$sample_data$sample_id)
print(obj_HH$data$tax_data$otu_id)
print(colnames(obj_HH$data$otu_table))

obj_HH$data$otu_props <- calc_obs_props(obj_HH, "otu_table", cols = obj_HH$data$sample_data$sample_id)

obj_HH$data$tax_abund <- calc_taxon_abund(obj_HH, "otu_props", cols = obj_HH$data$sample_data$sample_id)

obj_HH$data$tax_mean_prop <- calc_group_mean(obj_HH, data = "tax_abund",
                                          cols = obj_HH$data$sample_data$sample_id,
                                          groups = "mean_prop")

set.seed(123)
obj_HH %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", "#eae2b7", 
                       ifelse(taxon_names == "Rickettsiales", "#118ab2",
                       ifelse(taxon_names == "Rhodobacterales", "#003049",
                       ifelse(taxon_names == "Pseudomonadales", "#287271",
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", "#ef476f",
                       ifelse(taxon_names == "Oligoflexales", "#d62828",
                       ifelse(taxon_names == "Mycoplasmatales", "#fcbf49",
                       ifelse(taxon_names == "Gracilibacteria", "#06d6a0",
                       ifelse(taxon_names == "Bacteroidales", "#83d483", 
                              "gray"))))))))),
            initial_layout = "re", layout = "da",
            title = "HH",
            output_file = "metacoder_HH_plot_order_RA.pdf")
```

```{r}
#Healthy HH samples
obj_DD = parse_phyloseq(ps_f2_DD_pruned)

print(obj_DD$data$sample_data$sample_id)
print(obj_DD$data$tax_data$otu_id)
print(colnames(obj_DD$data$otu_table))

obj_DD$data$otu_props <- calc_obs_props(obj_DD, "otu_table", cols = obj_DD$data$sample_data$sample_id)

obj_DD$data$tax_abund <- calc_taxon_abund(obj_DD, "otu_props", cols = obj_DD$data$sample_data$sample_id)

obj_DD$data$tax_mean_prop <- calc_group_mean(obj_DD, data = "tax_abund",
                                          cols = obj_DD$data$sample_data$sample_id,
                                          groups = "mean_prop")
cols <- c("Rhodobacterales"="#003049", "Oligoflexales"="#d62828", "Flavobacteriales" = "#f77f00", "Mycoplasmatales" = "#fcbf49", "Verrucomicrobiales" = "#eae2b7", "Pseudomonadales" = "#287271", "Peptostreptococcales-Tissierellales" = "#ef476f", "Rickettsiales" = "#118ab2", "Gracilibacteria" = "#06d6a0", "Bacteroidales" = "#83d483")

set.seed(123)
obj_DD %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", "#eae2b7", 
                       ifelse(taxon_names == "Rickettsiales", "#118ab2",
                       ifelse(taxon_names == "Rhodobacterales", "#003049",
                       ifelse(taxon_names == "Pseudomonadales", "#287271",
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", "#ef476f",
                       ifelse(taxon_names == "Oligoflexales", "#d62828",
                       ifelse(taxon_names == "Mycoplasmatales", "#fcbf49",
                       ifelse(taxon_names == "Gracilibacteria", "#06d6a0",
                       ifelse(taxon_names == "Bacteroidales", "#83d483", 
                              "gray"))))))))),
            initial_layout = "re", layout = "da",
            title = "DD",
            output_file = "metacoder_DD_plot_order_RA.pdf")
```

```{r}
set.seed(123)

taxa_to_keep <- c("Verrucomicrobiales", "Verrucomicrobiae", "Verrucomicrobiota", "Rickettsiales", "Rhodobacterales", "Alphaproteobacteria","Proteobacteria", "Pseudomonadales", "Gammaproteobacteria", "Peptostreptococcales-Tissierellales","Clostridia" , "Firmicutes", "Oligoflexales", "Oligoflexia", "Bdellovibrionota", "Mycoplasmatales", "Gracilibacteria")

order_to_hide <- "Gammaproteobacteria" 

obj_HH %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = ifelse((taxon_names %in% taxa_to_keep) & !(taxon_ranks == "Order" & taxon_names == order_to_hide), taxon_names, ""),
            node_size = mean_prop,
            make_node_legend = FALSE,
            node_size_digits = 1,
            node_size_axis_label = "Mean Abundance",
            node_size_range = c(0.013, 0.08),
            node_label_size_range = c(0.02, 0.04), 
            repel_labels = TRUE,
            overlap_avoidance = 4 ,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiae", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiota", "#eae2b7",
                       ifelse(taxon_names == "Rickettsiales", "#118ab2",
                       ifelse(taxon_names == "Rhodobacterales", "#caf0f8",
                       ifelse(taxon_names == "Alphaproteobacteria" & taxon_ranks=="Class", "#caf0f8",
                       ifelse(taxon_names == "Proteobacteria", "#caf0f8",
                       ifelse(taxon_names == "Pseudomonadales", "#287271",
                       ifelse(taxon_names == "Gammaproteobacteria" & taxon_ranks=="Class", "#287271",
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", "#ef476f",
                       ifelse(taxon_names == "Clostridia" & taxon_ranks=="Class", "#ef476f",
                       ifelse(taxon_names == "Firmicutes", "#ef476f",
                       ifelse(taxon_names == "Oligoflexales", "#d62828",
                       ifelse(taxon_names == "Oligoflexia", "#d62828",
                       ifelse(taxon_names == "Bdellovibrionota", "#d62828",
                       ifelse(taxon_names == "Mycoplasmatales", "#fcbf49",
                       ifelse(taxon_names == "Gracilibacteria", "#06d6a0",
                       ifelse(taxon_names == "Bacteroidales", "#83d483", 
                              "gray")))))))))))))))))),
            initial_layout = "re", layout = "da",
            title = "Healthy Tissue",
            output_file = "metacoder_HH_plot_order_RA_colored.pdf")
```

```{r}
set.seed(123)

taxa_to_label <- c("Verrucomicrobiales", "Verrucomicrobiae", "Verrucomicrobiota", "Rickettsiales", "Rhodobacterales", "Alphaproteobacteria","Proteobacteria", "Pseudomonadales", "Gammaproteobacteria", "Peptostreptococcales-Tissierellales","Clostridia" , "Firmicutes", "Oligoflexales", "Oligoflexia", "Bdellovibrionota", "Mycoplasmatales", "Gracilibacteria", "Bacteroidales", "Bacteroidota", "Bacteroidia")

gamma_hide <- c("Gammaproteobacteria")
alpha_hide<- "Alphaproteobacteria"
clos_hide<- "Clostridia"

obj_DD %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = ifelse((taxon_names %in% taxa_to_label) & !(taxon_ranks == "Order" & (taxon_names == gamma_hide | taxon_names == alpha_hide | taxon_names == clos_hide)), taxon_names, ""),
            node_size = mean_prop,
            make_node_legend = FALSE,
            node_size_digits = 1,
            node_size_axis_label = "Mean Abundance",
            node_size_range = c(0.013, 0.08),
            node_label_size_range = c(0.02, 0.04), 
            repel_labels = TRUE,
            overlap_avoidance = 4 ,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiae", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiota", "#eae2b7",
                       ifelse(taxon_names == "Rickettsiales", "#118ab2",
                       ifelse(taxon_names == "Rhodobacterales", "#caf0f8",
                       ifelse(taxon_names == "Alphaproteobacteria" & taxon_ranks=="Class", "#caf0f8",
                       ifelse(taxon_names == "Proteobacteria", "#caf0f8",
                       ifelse(taxon_names == "Pseudomonadales", "#287271",
                       ifelse(taxon_names == "Gammaproteobacteria" & taxon_ranks=="Class", "#287271",
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", "#ef476f",
                       ifelse(taxon_names == "Clostridia" & taxon_ranks=="Class", "#ef476f",
                       ifelse(taxon_names == "Firmicutes", "#ef476f",
                       ifelse(taxon_names == "Oligoflexales", "#d62828",
                       ifelse(taxon_names == "Oligoflexia", "#d62828",
                       ifelse(taxon_names == "Bdellovibrionota", "#d62828",
                       ifelse(taxon_names == "Mycoplasmatales", "#fcbf49",
                       ifelse(taxon_names == "Gracilibacteria", "#06d6a0",
                       ifelse(taxon_names == "Bacteroidales", "#83d483",
                       ifelse(taxon_names == "Bacteroidota", "#83d483",
                       ifelse(taxon_names == "Bacteroidia", "#83d483",
                              "gray")))))))))))))))))))),
            initial_layout = "re", layout = "da",
            title = "Disease Lesion",
            output_file = "metacoder_DD_plot_order_RA_colored.pdf")
```

Now make combined plot.

```{r}
hh_plot<- obj_HH %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = ifelse((taxon_names %in% taxa_to_keep) & !(taxon_ranks == "Order" & taxon_names == order_to_hide), taxon_names, ""),
            node_size = mean_prop,
            make_node_legend = FALSE,
            node_size_digits = 1,
            node_size_axis_label = "Mean Abundance",
            node_size_range = c(0.013, 0.08),
            node_label_size_range = c(0.03, 0.05), 
            repel_labels = TRUE,
            overlap_avoidance = 3,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiae", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiota", "#eae2b7",
                       ifelse(taxon_names == "Rickettsiales", "#118ab2",
                       ifelse(taxon_names == "Rhodobacterales", "#caf0f8",
                       ifelse(taxon_names == "Alphaproteobacteria" & taxon_ranks=="Class", "#caf0f8",
                       ifelse(taxon_names == "Proteobacteria", "#caf0f8",
                       ifelse(taxon_names == "Pseudomonadales", "#287271",
                       ifelse(taxon_names == "Gammaproteobacteria" & taxon_ranks=="Class", "#287271",
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", "#ef476f",
                       ifelse(taxon_names == "Clostridia" & taxon_ranks=="Class", "#ef476f",
                       ifelse(taxon_names == "Firmicutes", "#ef476f",
                       ifelse(taxon_names == "Oligoflexales", "#d62828",
                       ifelse(taxon_names == "Oligoflexia", "#d62828",
                       ifelse(taxon_names == "Bdellovibrionota", "#d62828",
                       ifelse(taxon_names == "Mycoplasmatales", "#fcbf49",
                       ifelse(taxon_names == "Gracilibacteria", "#06d6a0",
                              "gray"))))))))))))))))),
            initial_layout = "re", layout = "da")

dd_plot<- obj_DD %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = ifelse((taxon_names %in% taxa_to_label) & !(taxon_ranks == "Order" & (taxon_names == gamma_hide | taxon_names == alpha_hide | taxon_names == clos_hide)), taxon_names, ""),
            node_size = mean_prop,
            make_node_legend = FALSE,
            node_size_digits = 1,
            node_size_axis_label = "Mean Abundance",
            node_size_range = c(0.013, 0.08),
            node_label_size_range = c(0.03, 0.05), 
            repel_labels = TRUE,
            overlap_avoidance = 5,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiae", "#eae2b7",
                       ifelse(taxon_names == "Verrucomicrobiota", "#eae2b7",
                       ifelse(taxon_names == "Rickettsiales", "#118ab2",
                       ifelse(taxon_names == "Rhodobacterales", "#caf0f8",
                       ifelse(taxon_names == "Alphaproteobacteria" & taxon_ranks=="Class", "#caf0f8",
                       ifelse(taxon_names == "Proteobacteria", "#caf0f8",
                       ifelse(taxon_names == "Pseudomonadales", "#287271",
                       ifelse(taxon_names == "Gammaproteobacteria" & taxon_ranks=="Class", "#287271",
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", "#ef476f",
                       ifelse(taxon_names == "Clostridia" & taxon_ranks=="Class", "#ef476f",
                       ifelse(taxon_names == "Firmicutes", "#ef476f",
                       ifelse(taxon_names == "Oligoflexales", "#d62828",
                       ifelse(taxon_names == "Oligoflexia", "#d62828",
                       ifelse(taxon_names == "Bdellovibrionota", "#d62828",
                       ifelse(taxon_names == "Mycoplasmatales", "#fcbf49",
                       ifelse(taxon_names == "Gracilibacteria", "#06d6a0",
                       ifelse(taxon_names == "Bacteroidales", "#83d483",
                       ifelse(taxon_names == "Bacteroidota", "#83d483",
                       ifelse(taxon_names == "Bacteroidia", "#83d483",
                              "gray")))))))))))))))))))),
            initial_layout = "re", layout = "da")

library(gridExtra)

grid.arrange(hh_plot, dd_plot, nrow = 1)  # Adjust nrow/ncol for layout

library(patchwork)

# Combine plots with annotations
combined_plot <- hh_plot + dd_plot 

print(combined_plot)

pdf("comparative_metacoder_trees.pdf", width = 12, height = 8)
combined_plot
dev.off()
```

Now to make them colored by relative abundance

```{r}
set.seed(123)
obj_HH %>% 
  filter_taxa(mean_prop > 0.002) %>%# subset to the order rank
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", mean_prop,
                       ifelse(taxon_names == "Verrucomicrobiae", mean_prop,
                       ifelse(taxon_names == "Verrucomicrobiota", mean_prop,
                       ifelse(taxon_names == "Rickettsiales", mean_prop,
                       ifelse(taxon_names == "Rhodobacterales", mean_prop,
                       ifelse(taxon_names == "Alphaproteobacteria" & taxon_ranks=="Class", mean_prop,
                       ifelse(taxon_names == "Proteobacteria", mean_prop,
                       ifelse(taxon_names == "Pseudomonadales", mean_prop,
                       ifelse(taxon_names == "Gammaproteobacteria" & taxon_ranks=="Class", mean_prop,
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", mean_prop,
                       ifelse(taxon_names == "Clostridia" & taxon_ranks=="Class", mean_prop,
                       ifelse(taxon_names == "Firmicutes", mean_prop,
                       ifelse(taxon_names == "Oligoflexales", mean_prop,
                       ifelse(taxon_names == "Oligoflexia", mean_prop,
                       ifelse(taxon_names == "Bdellovibrionota", mean_prop,
                       ifelse(taxon_names == "Mycoplasmatales", mean_prop,
                       ifelse(taxon_names == "Gracilibacteria", mean_prop,
                              "gray"))))))))))))))))),
            initial_layout = "re", layout = "da",
            title = "HH",
            output_file = "metacoder_HH_plot_order_RA_coloredRA.pdf")
```

```{r}
set.seed(123)
obj_DD %>% 
  filter_taxa(mean_prop > 0.002) %>%
  filter_taxa(taxon_ranks == "Order", supertaxa = TRUE) %>%
  heat_tree(node_label = taxon_names,
            node_size = mean_prop,
            make_node_legend = TRUE,
            node_color = ifelse(taxon_names == "Verrucomicrobiales", mean_prop,
                       ifelse(taxon_names == "Verrucomicrobiae", mean_prop,
                       ifelse(taxon_names == "Verrucomicrobiota", mean_prop,
                       ifelse(taxon_names == "Rickettsiales", mean_prop,
                       ifelse(taxon_names == "Rhodobacterales", mean_prop,
                       ifelse(taxon_names == "Alphaproteobacteria" & taxon_ranks=="Class", mean_prop,
                       ifelse(taxon_names == "Proteobacteria", mean_prop,
                       ifelse(taxon_names == "Pseudomonadales", mean_prop,
                       ifelse(taxon_names == "Gammaproteobacteria" & taxon_ranks=="Class", mean_prop,
                       ifelse(taxon_names == "Peptostreptococcales-Tissierellales", mean_prop,
                       ifelse(taxon_names == "Clostridia" & taxon_ranks=="Class", mean_prop,
                       ifelse(taxon_names == "Firmicutes", mean_prop,
                       ifelse(taxon_names == "Oligoflexales", mean_prop,
                       ifelse(taxon_names == "Oligoflexia", mean_prop,
                       ifelse(taxon_names == "Bdellovibrionota", mean_prop,
                       ifelse(taxon_names == "Mycoplasmatales", mean_prop,
                       ifelse(taxon_names == "Gracilibacteria", mean_prop,
                       ifelse(taxon_names == "Bacteroidales", mean_prop,
                       ifelse(taxon_names == "Bacteroidia", mean_prop,
                       ifelse(taxon_names == "Bacteroidota", mean_prop,
                              "gray")))))))))))))))))))),
            initial_layout = "re", layout = "da",
            title = "DD",
            output_file = "metacoder_DD_plot_order_RA_colored.pdf")
```

```{r}
ps_f2_ra #183 ASVs in 132 samples

ps_f2_ra_genus<- tax_glom(ps_f2_ra, taxrank = "Genus")
ps_f2_ra_genus_df <- as.data.frame(t(otu_table(ps_f2_ra_genus)))
mean_abund <- rowMeans(ps_f2_ra_genus_df)
top_15 <- names(sort(mean_abund, decreasing = TRUE))[1:15]
ps_top15_genus <- prune_taxa(top_15, ps_f2_ra_genus)
melted_data <- psmelt(ps_top15_genus)

#Assign colors per phylum
length(unique(melted_data$Phylum))
unique(melted_data$Phylum)
unique(melted_data$Genus)

cols <- c("Bacteroidota"="#2A9D8F","Cyanobacteria"="#264653","Proteobacteria"="#E9C46A",
          "Firmicutes"="#F4A261", "Bdellovibrionota" = "#9a031e")
library(stringr)

pdf("Bubbleplot_Top15_Genera.pdf",width=20,height=8.5)
bubplot <- ggplot(melted_data, aes(x=Sample, y=Genus, size=Abundance, fill=Phylum))+
  geom_point(shape=21)+
  scale_size(range = c(0, 17), name="Relative Abundance")+
  theme(axis.text.x = element_blank())+
  facet_grid(.~fct_relevel(Condition, "HH", "HD", "DD"), scales = "free_x", space = "free_x")+
  scale_fill_manual(values = cols, labels = c("Bacteroidota", "Bdellovibrionota", "Cyanobacteria", "Firmicutes", "Proteobacteria")) +
  theme(strip.background = element_blank())+
  xlab("Coral") +
  ylab("Lowest Taxonomic Assignment") +
  theme(panel.grid.major=element_line(linetype=1,color="grey"),
        panel.background = element_blank()) +
  theme(axis.text.x = element_blank())+
  theme(axis.title.x = element_text(size = 22, face = "bold")) +
  theme(strip.background =element_rect(fill="white"))+
  theme(strip.text = element_text(face = "bold"))+
  theme(strip.text = element_text(size = 22))+
  theme(axis.title.y = element_text(size = 22, face = "bold"))+
  theme(axis.title.x = element_text(size = 22, face = "bold"))+
  theme(axis.text.y = element_text(size=18, color = "black"))+
  theme(legend.title = element_text(size = 18, face = "bold"))+
  theme(legend.text = element_text(size = 18))+
  guides(fill = guide_legend(override.aes = list(size=7.5, vjust=30.5))) +
  theme(panel.border = element_blank()) +
  theme(panel.spacing = unit(1, "lines")) +
  theme(strip.background = element_blank())

bubplot
dev.off()
```

We were asked by folks in the FLP group to look at the functional profiles of the microbes and really focus on the nitrogen cycling bacteria. So we are going to run the ASVs through PICRUST2 on the HiperGator. In order to do this, we need to transpose the OTU table and then write it out as a tab-delimited file.

```{r}
ps_f2_otu_table<- t(otu_ps_f2)
rownames(ps_f2_otu_table) <- paste0("ASV", seq_len(nrow(ps_f2_otu_table)))
write.table(ps_f2_otu_table, "ps_f2_otu_table.txt", sep="\t",col.names=TRUE)

ps_f2_otu_table<- as.data.frame(ps_f2_otu_table)
library(readr)
write.table(ps_f2_otu_table, "ps_f2_otu_table.tsv", sep = "\t", row.names = TRUE, quote = FALSE)
```

##DESeq Diseased Corals with and Without SCTLD History

```{r}
ps_f2
ps_fs_sctld_flp<- subset_samples(ps_f2, Condition=="DD")
ps_f2_sctld_flp<- subset_samples(ps_fs_sctld_flp, Disease_History == "SCTLD" | Disease_History == "FLP_only")

ps_f2_sctld_flp

ps_f2_sctld_flp_pruned <- prune_taxa(taxa_sums(ps_f2_sctld_flp) > 0, ps_f2_sctld_flp)
ps_f2_sctld_flp_pruned

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_f2_sctld_flp_pruned)$Disease_History<-factor(sample_data(ps_f2_sctld_flp_pruned)$Disease_History,levels=c("SCTLD","FLP_only"))
head(sample_data(ps_f2_sctld_flp_pruned)$Disease_History, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_f2_sctld_flp_pruned, ~ Disease_History)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results_SCTLD_FLP_Only.txt",sep="\t",col.names=NA)
sig_dif<- read.table("DESeq2_results_SCTLD_FLP_Only.txt", header = TRUE)
colnames(sig_dif)[1]<- "ASV"

#Count how many orders
orders<-unique(sig_dif$Order)
print(orders)

order_cols<-c("Rhodospirillales"="#f47068", "Bacteroidales" = "#ffb3ae"  ,            "Pseudomonadales"  = "#fff4f1",  "Verrucomicrobiales" = "#1697a6",        
"Absconditabacteriales (SR1)"= "#0e606b","Rhizobiales" = "#ffc24b")  


#Create Figure for Significantly Different Taxa
pdf("Figure_DETaxa_SCTLD_FLP.pdf",width=11,height=8)
p1 <- ggplot(sig_dif,aes(log2FoldChange,Genus,fill=Order))+
  geom_point(size=5, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=order_cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Lowest Taxonomic Assignment")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank())
p1
dev.off()
```

```{r}
ps_f2_sctld_flp_pruned_ra<- transform_sample_counts(ps_f2_sctld_flp_pruned, function(x) x / sum(x))
sctld_flp_melt <- psmelt(ps_f2_sctld_flp_pruned_ra)

flp_vs_sctld_asvs<- sig_dif$ASV

filtered_phyloseq <- prune_taxa(flp_vs_sctld_asvs, ps_f2_sctld_flp_pruned_ra)
filtered_phyloseq <- prune_taxa(taxa_sums(filtered_phyloseq) > 0, filtered_phyloseq)

filtered_phyloseq

sctld_flp_melt <- psmelt(filtered_phyloseq)

##Make plot

sctld_flp_1<- sctld_flp_melt[sctld_flp_melt$OTU == flp_vs_sctld_asvs[1],]
print(flp_vs_sctld_asvs[1])

custom_labels <- c("TACGAAGGGAGCAAGCGTTGTTCGGAATTACTGGGCGTAAAGGGCGTGTAGGCGGATTATCAAGTTAGGGGTGAAATCCCGGGGCTCAACCTCGGCACTGCCTTTAAGACTGATAATCTAGAGTATGTGAGGGGATGATGGAATACCTAGTGTAGAGGTGAAATTCGTAGATATTAGGTGGAACACCAGTGGCGAAGGCGATTATCTGGCACATTACTGACGCTGAGGCGCGAAAGCGTGGGGAGCAAACGGG" = "Terasakiellaceae")

dd_plot_flp_sctld_1<- ggplot(sctld_flp_1, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_1
```

```{r}
##Make plot

sctld_flp_2<- sctld_flp_melt[sctld_flp_melt$OTU == flp_vs_sctld_asvs[2],]
print(flp_vs_sctld_asvs[2])

custom_labels <- c("TACGGAGGGTGTAAGCGTTATCCGGATTTATTGGGTTTAAAGGGTACGTAGGCGGTAATATAAGTCAGTGGTGAAATCCTGCAGCTCAACTGTAGAACTGCCATTGAAACTGTATTACTTGAGTATACTTGAGGTAGGCGGAATGAGTAGTGTAGCGGTGAAATGCTTAGATATTACTCAGAACACCGATTGCGAAGGCAGCTTACTAAGGTATAACTGACGCTGATGTACGAAAGCGTGGGGAGCGAACAGG" = "Roseimarinus")

dd_plot_flp_sctld_2<- ggplot(sctld_flp_2, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_2
```

```{r}
##Make plot

sctld_flp_3<- sctld_flp_melt[sctld_flp_melt$OTU == flp_vs_sctld_asvs[3],]
print(flp_vs_sctld_asvs[3])

custom_labels <- c("TACGGAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGTTTGTTAAGTCGGATGTGAAAGCCCCGGGCTCAACCTGGGAACTGCATCCGATACTGGCCGACTAGAGTACGAGAGAGGGAGGTAGAATTCCACGTGTAGCGGTGAAATGCGTAGATATGTGGAGGAATACCAGTGGCGAAGGCGGCCTCCTGGCTCGATACTGACGCTGAGGTGCGAAAGCGTGGGGAGCAAACAGG" = "Halieaceae")

dd_plot_flp_sctld_3<- ggplot(sctld_flp_3, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_3
```

```{r}
sctld_flp_4<- sctld_flp_melt[sctld_flp_melt$OTU == flp_vs_sctld_asvs[4],]
print(flp_vs_sctld_asvs[4])

custom_labels <- c("TACGAAGGTCCCAAGCGTTGTTCGGAATCACTGGGCGTAAAGCGTATGTAGGCGGCGTGGAAAGTCAGGTGTGAAAGCCCGGGGCTCAACCCCGGAATTGCACCCGATACTCCCACGCTAGAGTAATGGAGGGGAGTCTGGAATTCACGGTGTAGCAGTGAAATGCGTGGATATCGTGAGGAACACTAGTGGCGAAGGCGAGACTCTGGACATTTACTGACGCTGAGATACGAAGGCTAGGGGAGCGAATGGG" = "Roseibacillus")

dd_plot_flp_sctld_4<- ggplot(sctld_flp_4, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_4
```

```{r}
sctld_flp_5<- sctld_flp_melt[sctld_flp_melt$OTU == flp_vs_sctld_asvs[5],]
print(flp_vs_sctld_asvs[5])

custom_labels <- c("TACGTAGGTCCCGAGCGTTATCCGGATTTACTGGGCGTAAAGCGTCCGTAGCTTGGTATGTAAGTCTATTTTTAAAGACTCAGATTCAATCAGAGGAAGGGAATAGATACTGCATATCCTAGAGGATCTTAGAGGTTAGTGGAATTTCCGGTGGAGTGGTGAAATGCGTTGATATCGGAAGGAACGCCGAAAGCGAAAGCAGCTAACTATAATTTTCCTGACAGTGATGGACGAAAGTTTGGGTAGCAAACGGG" = "Absconditabacteriales (SR1)")

dd_plot_flp_sctld_5<- ggplot(sctld_flp_5, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_5
```

```{r}
sctld_flp_6<- sctld_flp_melt[sctld_flp_melt$OTU == flp_vs_sctld_asvs[6],]
print(flp_vs_sctld_asvs[6])

custom_labels <- c("TACGAAGGGGGCTAGCGTTGTTCGGAATCACTGGGCGTAAAGCGCGCGTAGGCGGACTTTTAAGTCAGGGGTGAAATCCCGAGGCTCAACCTCGGAACTGCCTTTGATACTGGAAGTCTTGAGTTCGAGAGAGGTGAGTGGAATACCGAGTGTAGAGGTGAAATTCGTAGATATTCGGTGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGCGTGGGGAGCAAACAGG" = "Cohaesibacter")

dd_plot_flp_sctld_6<- ggplot(sctld_flp_6, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_6
```

```{r}
combined_plot<- plot_grid(dd_plot_flp_sctld_1, dd_plot_flp_sctld_2, dd_plot_flp_sctld_3, dd_plot_flp_sctld_4,dd_plot_flp_sctld_5,dd_plot_flp_sctld_6, ncol=3, nrow = 2, align="v") +
  theme(plot.margin = margin(b = 15, l=15, unit = "pt"))

combined_plot

final_plot <- ggdraw(combined_plot) +
  draw_label("Health History", x = 0.5, y=0.0, size = 14, vjust = -0.5) +
  draw_label("Relative Abundance", x = 0, y = 0.5, angle = 90, size = 14, vjust = 1.5)

final_plot

pdf("combined_relative_abundance_plots_FLP_vs_SCTLD.pdf", width=8, height=6)
final_plot
dev.off()
```

#ASVs in DD corals not in HH corals

```{r}
ps_fs_dd<- subset_samples(ps_f2, Condition=="DD")
ps_f2_DD <- prune_taxa(taxa_sums(ps_fs_dd) > 0, ps_fs_dd)
ps_f2_DD

ps_f2_dd_100 <- phyloseq_filter_prevalence(
  physeq = ps_f2_DD,
  prev.trh = 1.0,  # 100% prevalence threshold
  abund.trh = NULL, # Optional abundance threshold
  threshold_condition = "OR" # Keep taxa meeting either threshold
)
ps_f2_dd_100

asv_vector <- taxa_names(ps_f2_dd_100)

dd_seqs<- names(sequences[sequences %in% asv_vector])

```

```{r}
ps_fs_hh<- subset_samples(ps_f2, Condition=="HH")
ps_f2_HH <- prune_taxa(taxa_sums(ps_fs_hh) > 0, ps_fs_hh)
ps_f2_HH

ps_f2_HH_asvs<- taxa_names(ps_f2_HH)

taxa_present <- asv_vector %in% ps_f2_HH_asvs

```

```{r}
ps_f2_melt<- psmelt(ps_f2)

ps_f2_melt_1<- ps_f2_melt[ps_f2_melt$OTU == asv_vector[1],]
print(asv_vector[1])

custom_labels <- c("TACGGGAGTGGCAAGCGTTATCCGGAATTATTGGGCGTAAAGCGTCCGCAGGCGGCCCTTCAAGTCTGCTGTTAAAAAGTGGAGCTTAACTCCATCATGGCAGTGGAAACTGTTGGGCTTGAGTGTGGTAGGGGCAGAGGGAATTCCCGGTGTAGCGGTGAAATGCGTAGATATCGGGAAGAACACCAGTGGCGAAGGCGCTCTGCTGGGCCATAACTGACGCTCATGGACGAAAGCCAGGGGAGCGAAAGGG" = "Cohaesibacter")

dd_plot_flp_sctld_6<- ggplot(sctld_flp_6, aes(x =Disease_History, y = Abundance, fill = Disease_History)) +
  geom_boxplot(alpha = 0.7,  outlier.shape=NA, aes(fill = Disease_History)) +
  geom_jitter(position=position_jitter(width=.1,height=0),aes(fill = Disease_History, colour="black"),size=4, pch=21, colour = "black")+
  labs(x = "Genus", y = "Relative Abundance", fill = "Disease_HIstory") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~OTU, labeller = labeller(OTU = custom_labels)) +
  theme_classic() +
  theme(panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black")) +
  theme(axis.text.x = element_text(color="black")) +
  theme(legend.position = "none") +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank())

dd_plot_flp_sctld_6
```

```{r}
ps_f2_melt_2<- ps_f2_melt[ps_f2_melt$OTU == asv_vector[2],]

ps_f2_melt_3<- ps_f2_melt[ps_f2_melt$OTU == asv_vector[3],]
```

### Unique to OATLD

We need to find which ASVs were enriched in Oatld and were indicators of oatld but didn't match SCTLD bioindicators/

```{r}
sctld_indicators<- tidy_blast_100_no_repeats

oatld_enriched<- sig_dif_hh_dd
oatld_indics<- dd_ind_asvs_w_tax
```

```{r}
#First, create a column of indicator species output that is the ASV sequence
oatld_indics<- rownames_to_column(oatld_indics, var = "ASV")

#Which microbes enriched in OATLD were identical to those in indicator species of lesions?
oatld_microbes<- oatld_enriched[oatld_enriched$ASV %in% oatld_indics$ASV, ]

#Which ones are not found in the BLASTn results of the SCTLD indicators?
oatld_microbes<- oatld_microbes %>% 
  filter(!ASV %in% sctld_indicators$sequence)

print(oatld_microbes$ASV[oatld_microbes$Genus=="Cribrihabitans"])
print(oatld_microbes$ASV[oatld_microbes$Genus=="Ferrimonas"])
print(oatld_microbes$ASV[oatld_microbes$Genus=="Mycoplasma"])
print(oatld_microbes$ASV[oatld_microbes$Genus=="Aurantivirga"])
print(oatld_microbes$ASV[oatld_microbes$Genus=="Micavibrionaceae"])

#Cribrihabitans
print(names(sequences)[which(sequences == "TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTCAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG")])

#Ferrimonas
print(names(sequences)[which(sequences == "TACGGAGGGTGCGAGCGTTAATCGGAATCACTGGGCGTAAAGCGCATGCAGGCGGTTTGCTAAGCGAGATGTGAAAGCCCCGGGCTTAACCTGGGAACTGCATTTCGAACTGGCGAACTAGAGTCTTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCAGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTCAGATGCGAAAGCGTGGGGAGCAAACAGG")])

#Mycoplasma
print(names(sequences)[which(sequences == "TACATAGGCCCCAAGCGTTATCCGGATTTACTGGGCGTAAAGAGTGCGCAGGTGATTTCACAAGTCTGTATTGAAAGGCTGGAGCTCAACTCCAGTTTGGTATGGAAACTGTGTTATTAGATTATGATAGAGGTTAGCGGAATTCCTAATGTAGCGGTGGAATGCGTAGATATTAGGAGGAACACCAAAAGCGAAGGCAGCTAACTGGGTCATTAATGACACTGGTGCACGAAAGCGTGGGGAGCAAACAGG")])

#Aurantivirga
print(names(sequences)[which(sequences == "TACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGTCAATTAAGTCAGAGGTGAAATCCCATAGCTTAACTATGGAACTGCCTTTGATACTGGTTGACTTGAGTTATACGGAAGTAGATAGAATAAGTAGTGTAGCGGTGAAATGCATAGATATTACTTAGAATACCGATTGCGAAGGCAGTCTACTACGTATATACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG")])

#Micavibrionaceae
print(names(sequences)[which(sequences == "TACTGGGGGAGCGAGCGTTGTTCGGAATCACTGGGCGTAAAGCGTGCGCAGGCGGTTTGGAAAGTCAGAAGTGAAATCCCAGAGCTCAACTCTGGAACTGCTTTTGAAACTTCCTTACTAGAGTTTGGTAGAGGTTGGGAGAATTCCTAGTGTAGAGGTGAAATTCGTAGATATTAGGAGGAATACCAGTGGCGTAGGCGCCCAACTGGACCATTACTGACGCTCAGGCACGAAAGCGTGGGGAGCAAACAGG")])

```

```{r}
library(cowplot)

combined_plot<- plot_grid(dd_plot_8, dd_plot_9, dd_plot_14, dd_plot_15, dd_plot_19, ncol=3, nrow = 2, align="v") +
  theme(plot.margin = margin(b = 15, l=15, unit = "pt"))

combined_plot

final_plot <- ggdraw(combined_plot) +
  draw_label("Health Condition", x = 0.5, y=0.0, size = 14, vjust = -0.5) +
  draw_label("Relative Abundance", x = 0, y = 0.5, angle = 90, size = 14, vjust = 1.5)

final_plot

pdf("combined_relative_abundance_plots_OATLD_unique_ASVs_updated.pdf", width=8, height=8)
final_plot
dev.off()
```

#CLAMtest

```{r}

###############################################
#CLAM Test PS
###############################################
otu_table <- as.data.frame(otu_table(ps_f2_HH_DD_pruned))
sample_data <- as.data.frame(sample_data(ps_f2_HH_DD_pruned))
# Assuming 'group' is a column in your sample data
env_var <- sample_data$Condition
clam_results_ps <- vegan::clamtest(otu_table, env_var, alpha = 0.05, specialization = 2/3)
print(clam_results_ps)

#Attach taxonomic data to ASVs
tax_table <- as.data.frame(tax_table(ps_f2_HH_DD_pruned))

tax_table$Species<- rownames(tax_table)
clam_ps_w_tax<- merge(clam_results_ps, tax_table, by = "Species")

###How many of each group?
#Subset by Generalists
clam_ps_w_tax_gen<- clam_ps_w_tax[clam_ps_w_tax$Classes=="Generalist",]
nrow(clam_ps_w_tax_gen)

#Subset by Healthy Specialists
clam_ps_w_tax_hh<- clam_ps_w_tax[clam_ps_w_tax$Classes=="Specialist_HH",]
nrow(clam_ps_w_tax_hh)

#Subset by Disease Specialists
clam_ps_w_tax_dd<- clam_ps_w_tax[clam_ps_w_tax$Classes=="Specialist_DD",]
nrow(clam_ps_w_tax_dd)

#Subset by Too Rare
clam_ps_w_tax_rare<- clam_ps_w_tax[clam_ps_w_tax$Classes=="Too_rare",]
nrow(clam_ps_w_tax_rare)

print(clam_results_ps)

summary(clam_results_ps)

c("DD" = "#d1495b", "HD" = "#edae49", "HH" = "#00798c")


colors <- c("Generalist" = "gray",
            "Specialist_HH" = "#00798c", 
            "Specialist_DD" = "#d1495b", 
            "Too_Rare" = "black")

labels<- c("Generalist", "Specialist Healthy", "Specialist Diseased", "Too Rare")

#sig_dif_hh_dd$ASV

x_1<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCACGTAGGCGGACTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTCCAAAACTATCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_1<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCACGTAGGCGGACTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTCCAAAACTATCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_2<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGAAAGGTGCAAGCGTTGTTCGGAATAACTGGGCGTAAAGCGCTCGTAGGCGGGAATTCAGGTCGATTGTGAAAGCCCAGAGCTCAACTTTGGAATTGCAGTCGAAACCGGATTTCTTGAATACTCGAGAGGGTGGGGGAATTCCTAGTGGAGAAGTGAAATTCGTAGAGATTAGGAGGAACACCAGAGGCGAAGGCGCCTACCTGGCGATGTATTGACGCTGAGGAGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_2<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGAAAGGTGCAAGCGTTGTTCGGAATAACTGGGCGTAAAGCGCTCGTAGGCGGGAATTCAGGTCGATTGTGAAAGCCCAGAGCTCAACTTTGGAATTGCAGTCGAAACCGGATTTCTTGAATACTCGAGAGGGTGGGGGAATTCCTAGTGGAGAAGTGAAATTCGTAGAGATTAGGAGGAACACCAGAGGCGAAGGCGCCTACCTGGCGATGTATTGACGCTGAGGAGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_3<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATCGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTCCAAAACTATCGGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_3<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATCGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTCCAAAACTATCGGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_4<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTAGGGGTGAAATCCCAGGGCTCAACCCTGGAACTGCCCTTAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_4<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTAGGGGTGAAATCCCAGGGCTCAACCCTGGAACTGCCCTTAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_5<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTCAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_5<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTCAAAACTCCCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_6<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTCCAAGCGTTATCCGGATTTATTGGGTTTAAAGGGTCCGTAGGCGGGGTTTTAAGTCAGTGGTGAAAGCCTACAGCTCAACTGTAGAACTGCCATTGAAACTGGAACTCTTGAATGTGATTGAGGTAGGCGGAATATGTCATGTAGCGGTGAAATGCTTAGATATGACATAGAACACCGATAGCGAAGGCAGCTTACCAAGTCATTATTGACGCTGATGGACGAAAGCGTGGGGAGCGAACAGG"] +1 
y_6<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTCCAAGCGTTATCCGGATTTATTGGGTTTAAAGGGTCCGTAGGCGGGGTTTTAAGTCAGTGGTGAAAGCCTACAGCTCAACTGTAGAACTGCCATTGAAACTGGAACTCTTGAATGTGATTGAGGTAGGCGGAATATGTCATGTAGCGGTGAAATGCTTAGATATGACATAGAACACCGATAGCGAAGGCAGCTTACCAAGTCATTATTGACGCTGATGGACGAAAGCGTGGGGAGCGAACAGG"] +1

x_7<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTGCAAGCGTTACTCGGAATCACTGGGCGTAAAGAGCGTGTAGGCGGGTTAATAAGTCAGATGTGAAAGCCTATGGCTCAACCATAGAACTGCATTTGAAACTGTTAACCTAGAATATGGGAGAGGTAGATGGAATTTCTGGTGTAGGGGTAAAATCCGTAGAGATCAGAAGGAATACCGATTGCGAAGGCGATCTACTGGAACATTATTGACGCTGAGACGCGAAAGCGTGGGGAGCAAACAGG"] +1 
y_7<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTGCAAGCGTTACTCGGAATCACTGGGCGTAAAGAGCGTGTAGGCGGGTTAATAAGTCAGATGTGAAAGCCTATGGCTCAACCATAGAACTGCATTTGAAACTGTTAACCTAGAATATGGGAGAGGTAGATGGAATTTCTGGTGTAGGGGTAAAATCCGTAGAGATCAGAAGGAATACCGATTGCGAAGGCGATCTACTGGAACATTATTGACGCTGAGACGCGAAAGCGTGGGGAGCAAACAGG"] +1

x_8<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTCCAAAACTATCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_8<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTGGAAAGTTGGGGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTCCAAAACTATCAGTCTAGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_9<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTAGTAAGTTAGAGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTTAATACTGCTAGTCTTGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1 
y_9<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGGGTTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGATTAGTAAGTTAGAGGTGAAATCCCGGGGCTCAACCCCGGAACTGCCTTTAATACTGCTAGTCTTGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGTGTGGGGAGCAAACAGG"] +1

x_10<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGTCAATTAAGTCAGAGGTGAAATCCTACAGCTTAACTGTAGAACTGCCTTTGATACTGGTTGACTTGAGTTATATGGAAGTAGGTAGAATGTGTAGTGTAGCGGTGAAATGCTTAGATATTACACAGAATACCGATTGCGAAGGCAGCCTACTACGTATATACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG"] +1 
y_10<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGTCAATTAAGTCAGAGGTGAAATCCTACAGCTTAACTGTAGAACTGCCTTTGATACTGGTTGACTTGAGTTATATGGAAGTAGGTAGAATGTGTAGTGTAGCGGTGAAATGCTTAGATATTACACAGAATACCGATTGCGAAGGCAGCCTACTACGTATATACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG"] +1

x_11<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGAGCTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGGCTGTTAAGTCAGAGGTGAAATCCCGGGGCTCAACCCCGGAACGGCCTTTGATACTGGCAGTCTGGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGCGTGGGGAGCAAACAGG"] +1 
y_11<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGAGCTAGCGTTGTTCGGAATTACTGGGCGTAAAGCGCGCGTAGGCGGGCTGTTAAGTCAGAGGTGAAATCCCGGGGCTCAACCCCGGAACGGCCTTTGATACTGGCAGTCTGGAGTTCGAGAGAGGTGAGTGGAATTCCGAGTGTAGAGGTGAAATTCGTAGATATTCGGAGGAACACCAGTGGCGAAGGCGGCTCACTGGCTCGATACTGACGCTGAGGTGCGAAAGCGTGGGGAGCAAACAGG"] +1

x_12<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTGCGAGCGTTAATCGGAATCACTGGGCGTAAAGCGCATGCAGGCGGTTTGCTAAGCGAGATGTGAAAGCCCCGGGCTTAACCTGGGAACTGCATTTCGAACTGGCGAACTAGAGTCTTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCAGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTCAGATGCGAAAGCGTGGGGAGCAAACAGG"] +1 
y_12<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTGCGAGCGTTAATCGGAATCACTGGGCGTAAAGCGCATGCAGGCGGTTTGCTAAGCGAGATGTGAAAGCCCCGGGCTTAACCTGGGAACTGCATTTCGAACTGGCGAACTAGAGTCTTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCAGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTCAGATGCGAAAGCGTGGGGAGCAAACAGG"] +1

x_13<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACATAGGCCCCAAGCGTTATCCGGATTTACTGGGCGTAAAGAGTGCGCAGGTGATTTCACAAGTCTGTATTGAAAGGCTGGAGCTCAACTCCAGTTTGGTATGGAAACTGTGTTATTAGATTATGATAGAGGTTAGCGGAATTCCTAATGTAGCGGTGGAATGCGTAGATATTAGGAGGAACACCAAAAGCGAAGGCAGCTAACTGGGTCATTAATGACACTGGTGCACGAAAGCGTGGGGAGCAAACAGG"] +1 
y_13<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACATAGGCCCCAAGCGTTATCCGGATTTACTGGGCGTAAAGAGTGCGCAGGTGATTTCACAAGTCTGTATTGAAAGGCTGGAGCTCAACTCCAGTTTGGTATGGAAACTGTGTTATTAGATTATGATAGAGGTTAGCGGAATTCCTAATGTAGCGGTGGAATGCGTAGATATTAGGAGGAACACCAAAAGCGAAGGCAGCTAACTGGGTCATTAATGACACTGGTGCACGAAAGCGTGGGGAGCAAACAGG"] +1

x_14<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGAAGGTCCCGAGCGTTATTCGGAATCACTGGGCGTAAAGGGAGCGTAGGCTGCGTGGTAAGTCAGATGTGAAATCTCAGGGCTCAACCCTGAAACTGCATCCGATACTGCCACGCTAGAGTAATGGAGGGGTAAGTGGAATTCTCAGTGTAGCAGTGAAATGCGTAGATATTGAGAGGAAGACCAACGGCGAAGGCAGCTTACTGGACATTTACTGACGCTGAGGCTCGAAGGCTAGGGTAGCGAAAGGG"] +1 
y_14<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGAAGGTCCCGAGCGTTATTCGGAATCACTGGGCGTAAAGGGAGCGTAGGCTGCGTGGTAAGTCAGATGTGAAATCTCAGGGCTCAACCCTGAAACTGCATCCGATACTGCCACGCTAGAGTAATGGAGGGGTAAGTGGAATTCTCAGTGTAGCAGTGAAATGCGTAGATATTGAGAGGAAGACCAACGGCGAAGGCAGCTTACTGGACATTTACTGACGCTGAGGCTCGAAGGCTAGGGTAGCGAAAGGG"] +1

x_15<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGCATGCAGGCGGTTACTCAAGTCAGATGTGAAATCCCCGAGCTTAACTCGGGAGGGTCATTTGAAACTGTGTAACTAGAGTCTTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTCAGATGCGAAAGCGTGGGTAGCAAACGGG"] +1 
y_15<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGCATGCAGGCGGTTACTCAAGTCAGATGTGAAATCCCCGAGCTTAACTCGGGAGGGTCATTTGAAACTGTGTAACTAGAGTCTTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTCAGATGCGAAAGCGTGGGTAGCAAACGGG"] +1

x_16<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGTCAATTAAGTCAGAGGTGAAATCCCATAGCTTAACTATGGAACTGCCTTTGATACTGGTTGACTTGAGTTATACGGAAGTAGATAGAATAAGTAGTGTAGCGGTGAAATGCATAGATATTACTTAGAATACCGATTGCGAAGGCAGTCTACTACGTATATACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG"] +1 
y_16<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTGCAAGCGTTATCCGGAATCATTGGGTTTAAAGGGTCCGCAGGCGGTCAATTAAGTCAGAGGTGAAATCCCATAGCTTAACTATGGAACTGCCTTTGATACTGGTTGACTTGAGTTATACGGAAGTAGATAGAATAAGTAGTGTAGCGGTGAAATGCATAGATATTACTTAGAATACCGATTGCGAAGGCAGTCTACTACGTATATACTGACGCTCATGGACGAAAGCGTGGGGAGCGAACAGG"] +1

x_17<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGGAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGCATGCAGGTGGTTCGTTAAGTCAGATGTGAAAGCCCGGGGCTCAACCTCGGAACTGCATTTGAAACTGGCGGACTAGAGTACTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACAGATACTGACACTCAGATGCGAAAGCGTGGGGAGCAAACAGG"] +1 
y_17<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGGAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGCATGCAGGTGGTTCGTTAAGTCAGATGTGAAAGCCCGGGGCTCAACCTCGGAACTGCATTTGAAACTGGCGGACTAGAGTACTGTAGAGGGGGGTAGAATTTCAGGTGTAGCGGTGAAATGCGTAGAGATCTGAAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACAGATACTGACACTCAGATGCGAAAGCGTGGGGAGCAAACAGG"] +1

x_18<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACGAAGGTCCCGAGCGTTATTCGGAATCACTGGGCGTAAAGGGAGCGTAGGCTGCGTGGTAAGTCAGATGTGAAAGCCCAGGGCTCAACCCTGGAACTGCATTCGATACTGCCATGCTAGAGTAATGGAGAGGTAGATGGAATTTTCGGTGTAGCAGTGAAATGCGTAGATATCGAAAGGAAGACCAATGGCGAAGGCAATCTACTGGACATTTACTGACGCTGAGGCTCGAAGGCTAGGGGAGCGAAAGGG"] +1 
y_18<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACGAAGGTCCCGAGCGTTATTCGGAATCACTGGGCGTAAAGGGAGCGTAGGCTGCGTGGTAAGTCAGATGTGAAAGCCCAGGGCTCAACCCTGGAACTGCATTCGATACTGCCATGCTAGAGTAATGGAGAGGTAGATGGAATTTTCGGTGTAGCAGTGAAATGCGTAGATATCGAAAGGAAGACCAATGGCGAAGGCAATCTACTGGACATTTACTGACGCTGAGGCTCGAAGGCTAGGGGAGCGAAAGGG"] +1

x_19<- clam_results_ps$Total_DD[clam_results_ps$Species=="TACTGGGGGAGCGAGCGTTGTTCGGAATCACTGGGCGTAAAGCGTGCGCAGGCGGTTTGGAAAGTCAGAAGTGAAATCCCAGAGCTCAACTCTGGAACTGCTTTTGAAACTTCCTTACTAGAGTTTGGTAGAGGTTGGGAGAATTCCTAGTGTAGAGGTGAAATTCGTAGATATTAGGAGGAATACCAGTGGCGTAGGCGCCCAACTGGACCATTACTGACGCTCAGGCACGAAAGCGTGGGGAGCAAACAGG"] +1 
y_19<- clam_results_ps$Total_HH[clam_results_ps$Species=="TACTGGGGGAGCGAGCGTTGTTCGGAATCACTGGGCGTAAAGCGTGCGCAGGCGGTTTGGAAAGTCAGAAGTGAAATCCCAGAGCTCAACTCTGGAACTGCTTTTGAAACTTCCTTACTAGAGTTTGGTAGAGGTTGGGAGAATTCCTAGTGTAGAGGTGAAATTCGTAGATATTAGGAGGAATACCAGTGGCGTAGGCGCCCAACTGGACCATTACTGACGCTCAGGCACGAAAGCGTGGGGAGCAAACAGG"] +1


```
```{r}
pdf("CLAM_plot.pdf", width = 8.25, height = 6)
plot(clam_results_ps, xlab = "DD (Abundance + 1)", ylab = "HH (Abundance + 1)", 
     pch = c(1,16,16,1), col=colors, col.lines = c("black", "black", "black"), 
     lty = c(2,2,2), cex=1.5, position = "topleft")
legend("topleft",legend = labels, pch = c(1,16,16,1), col=colors)
text(x=x_1, y=y_1, labels = "Ruegeria", pos=1)
text(x=x_2, y=y_2, labels = "Pseudobacteriovorax", pos=2)
text(x=x_3,y=y_3, labels = "Ruegeria", pos = 1)
text(x=x_4,y=y_4, labels = "Rhodobacteraceae", pos = 4)
text(x=x_5,y=y_5, labels = "Cribrihabitans", pos=2)
text(x=x_6,y=y_6, labels = "Cryomorphaceae", pos=2)
text(x=x_7,y=y_7, labels = "Halarcobacter", pos=2)
text(x=x_8,y=y_8, labels = "Rhodobacteraceae", pos=3)
text(x=x_9,y=y_9, labels = "Thalassobius", pos=2)
text(x=x_10,y=y_10, labels = "Tenacibaculum", pos=2)
text(x=x_11,y=y_11, labels = "Rhodobacteraceae", pos=2)
text(x=x_12,y=y_12, labels = "Ferrimonas", pos=4)
text(x=x_13,y=y_13, labels = "Mycoplasma", pos=2)
text(x=x_14,y=y_14, labels = "Rubritalea", pos=2)
text(x=x_15,y=y_15, labels = "Vibrionaceae", pos=2)
text(x=x_16,y=y_16, labels = "Aurantivirga", pos=2)
text(x=x_17,y=y_17, labels = "Vibrio", pos=2)
text(x=x_18,y=y_18, labels = "Rubritalea", pos=2)
text(x=x_19,y=y_19, labels = "Micavibrionaceae", pos=3)
dev.off()
```

For the rest of the OATLD group, we need to get the relative abundance of each of the groups for each sample

```{r}
ps_f2_ra

#Enterobacterales
ps_enterobacterales <- subset_taxa(ps_f2_ra, Order == "Enterobacterales")
View(ps_enterobacterales@otu_table)
ps_enterobacterales_ras<- as.data.frame(ps_enterobacterales@otu_table)
ps_enterobacterales_ras$RA_Total <- rowSums(ps_enterobacterales_ras)

enterobacterales_ras <- data.frame(
  RowNames = rownames(ps_enterobacterales_ras),
  LastColumn = ps_enterobacterales_ras[, 42]
)

#Flavobacteriales
ps_Flavobacteriales <- subset_taxa(ps_f2_ra, Order == "Flavobacteriales")
View(ps_Flavobacteriales@otu_table)
ps_Flavobacteriales_ras<- as.data.frame(ps_Flavobacteriales@otu_table)
ps_Flavobacteriales_ras$RA_Total <- rowSums(ps_Flavobacteriales_ras)

Flavobacteriales_ras <- data.frame(
  RowNames = rownames(ps_Flavobacteriales_ras),
  LastColumn = ps_Flavobacteriales_ras[, 137]
)
Flavobacteriales_ras$LastColumn <- format(Flavobacteriales_ras, scientific = FALSE)

#Rhodobacterales
ps_Rhodobacterales <- subset_taxa(ps_f2_ra, Order == "Rhodobacterales")
View(ps_Rhodobacterales@otu_table)
ps_Rhodobacterales_ras<- as.data.frame(ps_Rhodobacterales@otu_table)
ps_Rhodobacterales_ras$RA_Total <- rowSums(ps_Rhodobacterales_ras)

Rhodobacterales_ras <- data.frame(
  RowNames = rownames(ps_Rhodobacterales_ras),
  LastColumn = ps_Rhodobacterales_ras[, 76]
)
Rhodobacterales_ras$LastColumn <- format(Rhodobacterales_ras, scientific = FALSE)

#Peptostreptococcales-Tessierellales
ps_Peptostreptococcales_Tessierelales <- subset_taxa(ps_f2_ra, Order == "Peptostreptococcales-Tissierellales")
View(ps_Peptostreptococcales_Tessierelales@otu_table)
ps_Peptostreptococcales_Tessierelales_ras<- as.data.frame(ps_Peptostreptococcales_Tessierelales@otu_table)
ps_Peptostreptococcales_Tessierelales_ras$RA_Total <- rowSums(ps_Peptostreptococcales_Tessierelales_ras)

Peptostreptococcales_Tessierelales_ras <- data.frame(
  RowNames = rownames(ps_Rhodobacterales_ras),
  LastColumn = ps_Rhodobacterales_ras[, 27]
)
Peptostreptococcales_Tessierelales_ras$LastColumn <- format(Peptostreptococcales_Tessierelales_ras, scientific = FALSE)

#Mycoplasmatales
ps_Mycoplasmatales <- subset_taxa(ps_f2_ra, Order == "Mycoplasmatales")
View(ps_Mycoplasmatales@otu_table)
ps_Mycoplasmatales_ras<- as.data.frame(ps_Mycoplasmatales@otu_table)
ps_Mycoplasmatales_ras$RA_Total <- rowSums(ps_Mycoplasmatales_ras)

Mycoplasmatales_ras <- data.frame(
  RowNames = rownames(ps_Mycoplasmatales_ras),
  LastColumn = ps_Mycoplasmatales_ras[, 1]
)
Mycoplasmatales_ras$LastColumn <- format(Mycoplasmatales_ras, scientific = FALSE)

ps_mycoplasmatales <- subset_taxa(ps_f2_ra, Order == "Mycoplasmatales")
```


#Make the Figures
```{r}
#Beta Dispersion
p2<-ggplot(dis.condition, aes(x=factor(Condition), y=distance)) +
  geom_boxplot(mapping =aes(x=Condition,y=distance, fill=Condition, color="black"), alpha = 0.6) +
  geom_point(position=position_jitter(width=.1, height=0),aes(colour="black",fill=Condition),size=4, pch=21, colour = "black") +
  scale_fill_manual(values = palette2) +
  scale_color_manual(values = c( "black")) +
  ylab("Distance to Centroid")+
  xlab("Health Condition")+
  guides(color="none") +
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(axis.text.x = element_text(colour="black"))+
  theme(axis.text.y = element_text(colour="black"))+
  theme(axis.text.x = element_text(size = 14))+
  theme(axis.text.y = element_text(size = 12))+
  theme(axis.title.x = element_text(size = 16))+
  theme(axis.title.y = element_text(size = 16))+
  theme(strip.text.x=element_text(size=16, face="bold"))+
  theme(legend.position = "none") +
  stat_compare_means(
    comparisons = list(c("DD", "HH")),
    method = "t.test",
    label = "p.signif") +
  scale_x_discrete(labels = c("DD" = "D", "HD" = "U", "HH" = "H")) 
p2

ggsave("beta_dispersion.png", plot = p2, width = 8, height = 6, dpi = 600)
```

```{r}
#Remake Combined PLot of Relative Abundance
library(cowplot)

combined_plot<- plot_grid(dd_plot_8, dd_plot_9, dd_plot_14, dd_plot_15, dd_plot_19, ncol=5, nrow = 1, align="v") +
  theme(plot.margin = margin(b = 30, l=30, unit = "pt"))

combined_plot

final_plot <- ggdraw(combined_plot) +
  draw_label("Health Condition", x = 0.5, y=0.0, size = 14, vjust = -1) +
  draw_label("Relative Abundance", x = 0, y = 0.5, angle = 90, size = 14, vjust = 1.5)

final_plot

pdf("combined_relative_abundance_plots_OATLD_unique_ASVs_updated.pdf", width=11, height=6)
final_plot
dev.off()
```
```{r}
#Differential ABundance Figure
ps_f2
ps_f2_HH_DD<- subset_samples(ps_f2, Condition == "HH" | Condition == "DD")
ps_f2_HH_DD

ps_f2_HH_DD_pruned <- prune_taxa(taxa_sums(ps_f2_HH_DD) > 0, ps_f2_HH_DD)
ps_f2_HH_DD_pruned

# Define the order of the conditions for testing
# In this order, the positive fold change values are increased in Diseased Tissue
sample_data(ps_f2_HH_DD_pruned)$Condition<-factor(sample_data(ps_f2_HH_DD_pruned)$Condition,levels=c("HH","DD"))
head(sample_data(ps_f2_HH_DD_pruned)$Condition, 10)

# DESEQ2 analysis on Dcyl Samples
dds = phyloseq_to_deseq2(ps_f2_HH_DD_pruned, ~ Condition)
dds
#filter rows with very few counts
dds <- dds[ rowSums(counts(dds)) > 5, ]
cts <- counts(dds)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds <- estimateSizeFactors(dds, geoMeans=geoMeans)
dds <- DESeq(dds,test="Wald", fitType="parametric")
res = results(dds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
dim(sigtab)
#save table of results
sig = as(sigtab, "matrix")
write.table(sig,"DESeq2_results.txt",sep="\t",col.names=NA)
sig_dif_hh_dd<- read.table("DESeq2_results.txt", header = TRUE)
colnames(sig_dif_hh_dd)[1]<- "ASV"

#Count how many orders
orders<-unique(sig_dif_hh_dd$Order)
print(orders)

cols <- c("Rhodobacterales"="#003049", "Oligoflexales"="#d62828", "Flavobacteriales" = "#f77f00", "Mycoplasmatales" = "#fcbf49", "Verrucomicrobiales" = "#eae2b7", "Enterobacterales" = "#c8553d", "Micavibrionales" = "#588b8b", "Campylobacterales" = "#bc6c25")


#Create Figure for Significantly Different Taxa
pdf("FigureDEtaxa.pdf",width=8,height=6)
p1 <- ggplot(sig_dif_hh_dd,aes(log2FoldChange,Genus,fill=Order))+
  geom_point(size=5, shape=21)+
  geom_errorbarh(aes(xmin=log2FoldChange-lfcSE, xmax=log2FoldChange+lfcSE),height=0.2)+
  scale_fill_manual(values=cols)+
  theme(plot.title = element_text(face="italic"))+
  labs(y= "Lowest Taxonomic Assignment")+
  theme(axis.title=element_text(size=14))+
  theme(legend.text=element_text(size=14))+
  theme(legend.title=element_text(size=16))+
  theme(axis.text=element_text(size=14, color="black"))+
  theme(legend.position = "right")+
  geom_vline(xintercept=0, linetype="dashed") +
  theme(panel.grid = element_blank()) +
  theme(panel.background = element_blank()) +
  scale_x_continuous(limits = c(-5, 30)) +
  scale_y_discrete(limits= c("Vibrionaceae", "Vibrio", "Thalassobius", "Tenacibaculum", "Ruegeria", "Rubritalea","Rhodobacteraceae", "Pseudobacteriovorax", "Mycoplasma","Micavibrionaceae", "Halarcobacter", "Ferrimonas", "Cryomorphaceae", "Cribrihabitans", "Aurantivirga"),
                   labels= c("Vibrionaceae", expression(italic("Vibrio")), expression(italic("Thalassobius")), 
                             expression(italic("Tenacibaculum")), 
                             expression(italic("Ruegeria")), expression(italic("Rubritalea")), "Rhodobacteraceae",
                             expression(italic("Pseudobacteriovorax")),
                             expression(italic("Mycoplasma")),
                             "Micavibrionaceae", expression(italic("Halarcobacter")), 
                             expression(italic("Ferrimonas")), "Cryomorphaceae",
                             expression(italic("Cribrihabitans")), 
                            expression(italic("Aurantivirga"))))
p1
dev.off()
```

